### 2024-06-16 code/eda-code/data_exploration_part_2.qmd
add previous scripts for historical purposes, refactoring to Python with PySpark implementation going forward

```{r}
library(readxl) #for loading Excel files
library(readr)
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(skimr) #for nice visualization of data 
library(here) #to set paths
library(tidyverse)
library(ggplot2)
library(tidycensus)
```

```{r}
getwd()
```

```{r}
# Import Data, Check Descriptive Statistics & Data Types
data_location <- here::here("starter-analysis-exercise","code","final_project","Electric_Vehicle_Population_Data.csv")
ev_data <- read.csv(data_location)
summary(ev_data)
dplyr::glimpse(ev_data)
```

```{r}
# Checking for duplicates:
duplicates <- duplicated(ev_data) # Check for duplicates
num_duplicates <- sum(duplicates) # Count of duplicates
print(num_duplicates)
```

```{r}
# Find total rows with MSRP prices
count_zero_msrp = length(which(ev_data$Base.MSRP == 0))
total_rows_count = nrow(ev_data)
total_price_data_points = total_rows_count - count_zero_msrp
print(paste0('Total rows: ', total_rows_count))
print(paste0('Total Missing MSRP Prices: ', count_zero_msrp))
print(paste0('Total Price Data Points: ', total_price_data_points))
```

```{r}
# Checking for Null values:
total_na <- sum(is.na(ev_data))
print(total_na)
```

```{r}
# Removing null values and outlier from MSRP:
ev_data_filtered <- na.omit(ev_data) # FIltered data with NA values removed
max_msrp_outlier <- ev_data_filtered %>%
  group_by(Electric.Vehicle.Type) %>%
  summarise(max = max(Base.MSRP)) 
max_msrp_outlier
filter_msrp_outlier <- ev_data_filtered %>%
  filter(Base.MSRP != 845000) %>%
  filter(Base.MSRP != 0)
filter_msrp_outlier
```

```{r}
# Create a boxplot of price distribution by electric vehicle type
p2_boxplot<-ggplot(filter_msrp_outlier, aes(x = Electric.Vehicle.Type, y = Base.MSRP)) +
  geom_boxplot() +
  labs(title = "Boxplot - Price Distribution vs. Electric Vehicle Type",
       x = "Electric Vehicle Type",
       y = "MSRP")
p2_boxplot 
```

```{r}
# Clean data for plot
price_by_state = ev_data_filtered %>% 
  group_by(Clean.Alternative.Fuel.Vehicle..CAFV..Eligibility) %>%
  filter(Base.MSRP != 0) %>%
  filter(Base.MSRP != 845000)

# Create a boxplot for price distribution by alternative fuel eligibility
p3_boxplot<-ggplot(price_by_state, aes(x = Clean.Alternative.Fuel.Vehicle..CAFV..Eligibility, y = Base.MSRP)) +
  geom_boxplot() +
  labs(title = "Boxplot - Price Distribution vs. Alt. Vehicle Eligibility",
       x = "Electric Vehicle Type",
       y = "MSRP")
p3_boxplot 
```

```{r}
# Clean data from plot
price_by_state = ev_data_filtered %>% 
  group_by(Clean.Alternative.Fuel.Vehicle..CAFV..Eligibility) %>%
  filter(Electric.Range != 0)

# Create a boxplot
p4_boxplot<-ggplot(price_by_state, aes(x = Clean.Alternative.Fuel.Vehicle..CAFV..Eligibility, y = Electric.Range)) +
  geom_boxplot() +
  labs(title = "Boxplot - Price Distribution vs. Electric Vehicle Type")
p4_boxplot 
```
```{r}
# Create a boxplot with the new categorical variable on the x-axis and height on the y-axis
p_boxplot<-ggplot(filter_msrp_outlier, aes(x = Make, y = Base.MSRP)) +
  geom_boxplot() +
  labs(title = "Boxplot - Price Distribution by Brand",
       x = "Brand",
       y = "Price")
p_boxplot
```
``` {r}
# Create a scatterplot with electric vehicle range growth over time

p_scatterplot<-ggplot(ev_data_filtered, aes(x = Model.Year, y = Electric.Range)) +
  geom_point() +
  labs(title = "Scatterplot - Electric Vehicle Range Growth over Time",
       x = "Year",
       y = "Electric Range")
plot(p_scatterplot)
# Save the scatterplot to a file
# ggsave("scatterplot.png", plot = p_scatterplot)
```

``` {r}
# Create a scatterplot - price vs. electric vehicle range correlation

p_scatterplot<-ggplot(filter_msrp_outlier, aes(x = Electric.Range, y = Base.MSRP)) +
  geom_point() +
  labs(title = "Scatterplot - Price vs. Range Correlation",
       x = "Vehicle Range",
       y = "MSRP")
plot(p_scatterplot)
```

```{r}
#For Electric.Range:
Q1 <- quantile(ev_data_filtered$Electric.Range , 0.25, na.rm = TRUE) # First Quartile
Q3 <- quantile(ev_data_filtered$Electric.Range , 0.75, na.rm = TRUE) # Third Quartile
IQR <- Q3 - Q1 # Compute Interquartile Range
lower_bound <- Q1 - 1.5 * IQR # Compute Lower Bound based on IQR
upper_bound <- Q3 + 1.5 * IQR # Compute Upper Bound based on IQR
upper_bound
```

```{r}
# Filter out outliers:
ev_data_Filtered_2 <- ev_data_filtered[ev_data_filtered$Electric.Range > lower_bound & ev_data_filtered$Electric.Range < upper_bound, ]
ev_data_Filtered_2
```

### 2024-06-16 code/processing-code/pre_process_tax_data.py

@@ -1,0 +1,24 @@
+from pandas import read_csv
+
+def convert_csv_to_parquet(year: int, dataset: str):
+    try:
+        print('Converting csv to parquet.')
+        csv_df = read_csv(f'{dataset}{year}.csv')
+        csv_df.to_parquet(f'{dataset}{year}.parquet')
+        print('Conversion successful!')
+    except Exception as e:
+        print('CSV to parquet conversion failed :', e)
+
+years = [
+    '2011','2012','2013',
+    '2014','2015','2016',
+    '2017','2018','2019',
+    '2020','2021',]
+
+datasets_to_convert = [
+    'tax_zip_code_',
+    'data_dict',]
+
+for year in years:
+    convert_csv_to_parquet(year=year, dataset=datasets_to_convert[0])
+    convert_csv_to_parquet(year=year, dataset=datasets_to_convert[1])

### 2024-06-16 products/manuscript/Manuscript.qmd
replace the template Manuscript with the one our team has been updating

---
title: "Example Manuscript Template for a Data Analysis Project"
subtitle: ""
author: Sri Lakshmi Chundru, Collin Real, Joaquin Ramirez, Seth Harris
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/apa.csl
---

The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. 
You don't need to have exactly these sections, but the content covering those sections should be addressed.

This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like html or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

-   https://www.google.com/publicdata/directory
-   https://www.data.gov/
-   https://github.com/awesomedata/awesome-public-datasets
-   https://registry.opendata.aws/
-   https://www.kaggle.com/datasets

- https://www.kaggle.com/datasets/jainaru/electric-vehicle-population

# Summary/Abstract
_Write a summary of your project._

The dataset obtained from Kaggle provides detailed information about Battery Electric Vehicles (BEVs) and Plug-in Hybrid Electric Vehicles (PHEVs) registered through the Washington State Department of Licensing (DOL). It comprises 181,458 observations and 17 variables, encompassing key details such as Vehicle Identification Number (VIN), county, city, model year, electric range, and more.

{{< pagebreak >}}

# Introduction

**- (a) Briefly describe what the data is, how it was collected, where you will get (or got) it from. How many observations do you have, what was measured? Anything else important to report about the data?**

The dataset obtained from Kaggle provides detailed information about Battery Electric Vehicles (BEVs) and Plug-in Hybrid Electric Vehicles (PHEVs) registered through the Washington State Department of Licensing (DOL). It comprises 181,458 observations and 17 variables, encompassing key details such as Vehicle Identification Number (VIN), county, city, model year, electric range, and more.

**- (b) At this stage you are not required to already have and show the data, but if you do, even better. Then add a few lines of code which load the data and using some of the commands you learned about, provide summary descriptions of the data.**
```{r}
data_location <- here::here("starter-analysis-exercise","code","final_project")
data <- read.csv("Electric_Vehicle_Population_Data.csv")
str(data)
head(data)
summary(data)
colSums(is.na(data))
```

**- (c) Explain the question you want to answer using the data. What will be your outcome(s) of interest (if any)? What (if any) specific predictors will you focus on? What relations/patterns are you looking for in the data?**

Question: What insights can be gleaned from the dataset regarding the electric vehicle landscape ?

## Objectives:

- Understand Geographic Distribution: Determine the distribution of electric vehicles across different counties and cities in Washington State.
- Identify Popular Models: Identify the most commonly registered electric vehicle makes and models.
- Analyze Electric Range: Investigate the distribution of electric range among registered electric vehicles.-
- Explore Clean Alternative Fuel Vehicle Eligibility: Examine the eligibility for clean alternative fuel vehicle incentives among different vehicle types.

## Outcomes of Interest
- Geographic Distribution: The distribution of electric vehicles across counties and cities in Washington State.
- Popular Models: The most frequently registered electric vehicle makes and models.
- Electric Range Analysis: The range distribution and average electric range of registered electric vehicles.
- Clean Alternative Fuel Vehicle Eligibility: The proportion of vehicles eligible for clean alternative fuel vehicle incentives.

## Predictors
- Geographic Variables: County, City, Postal Code.
- Vehicle Attributes: Model Year, Make, Model, Electric Vehicle Type, Electric Range.

## Relations/Patterns
- Geographic Distribution: Analyzing clusters of electric vehicles in specific regions.
- Popular Models: Identifying dominant electric vehicle makes and models.
- Electric Range Analysis: Understanding the distribution of electric ranges and potential implications for consumer preferences.
- Clean Alternative Fuel Vehicle Eligibility: Examining eligibility patterns and potential factors influencing eligibility.
- Correlations: Investigating any associations between electric vehicle characteristics and legislative districts or census tracts, which may offer insights into policy implications and consumer behavior.

**(d) As much as you know, suggest how you will analyze it. At this stage in the course, we haven’t covered analysis approaches yet, so you can keep things vague and non-technical here.**

- Descriptive Analysis: Summarize the dataset to understand its contents, including summary statistics and frequency tables.

-Visualization: Use plots and charts to visualize aspects like geographic distribution, electric range, and vehicle types.

- Comparative Analysis: Compare different groups within the dataset, such as BEVs vs PHEVs or eligibility for incentives by vehicle type.

- Correlation Analysis: Explore potential relationships between variables, such as electric range and base MSRP.

- Geospatial Analysis (optional): Use R packages to analyze spatial patterns in the data.

- Exploratory Data Analysis (EDA): Engage in exploratory analysis to identify interesting patterns or trends.

**(e) You are allowed, but not yet required, to provide background information for the question you plan to answer. For instance you can describe why it’s an interesting question, who else has done similar analyses, how your analysis will be new/different, etc. Similar to what you read in an introduction to a research paper. For the final report, you’ll need these parts. For part 1, they are not required, but you are welcome to already write down some of that. **

## Primary Research Question: What insights can be gleaned from the dataset regarding the electric vehicle landscape in Washington State?

Sub-questions:

-  How are electric vehicles distributed across different counties, cities, and States?
-  Which EV makes and models are most popular?
-  What is the range of electric vehicles in terms of electric range and base MSRP?
-  How does eligibility for clean alternative fuel vehicle incentives vary among different vehicle types?
-  Are there any significant correlations between EV characteristics and geographic or legislative factors?

**(d) Eventually, for your final report, what you write for this part will go into different sections of the full report. Some will go into the introduction, some in the methods section. You can already place these items there, or for now just write them as a single section.**

**Introduction**
Electric vehicles (EVs), including Battery Electric Vehicles (BEVs) and Plug-in Hybrid Electric Vehicles (PHEVs), are crucial in the global effort to reduce greenhouse gas emissions and dependence on fossil fuels. This dataset, sourced from the Washington State Department of Licensing (DOL) and obtained from Kaggle, provides a comprehensive overview of registered electric vehicles in Washington State. Analyzing this data offers valuable insights into the adoption patterns and characteristics of EVs, informing policies and infrastructure development to further promote sustainable transportation.

**Objectives:**

**Geographic Distribution:** Determine the distribution of electric vehicles across different counties and cities.
**Popular Models:** Identify the most commonly registered EV makes and models.
**Electric Range Analysis:** Investigate the distribution of electric range among registered EVs.
**CAFV Eligibility:** Examine the eligibility for clean alternative fuel vehicle incentives among different vehicle types.
**Correlation Analysis:** Explore potential correlations between EV characteristics and geographic or legislative factors.

**Methods**

**Data Cleaning and Preparation:**

- Address missing values and ensure correct data types.
- Transform variables as necessary for analysis.

To ensure the dataset is ready for analysis, the following steps were undertaken:

**1. Data Import and Initial Inspection:**

- The dataset Electric_Vehicle_Population_Data.csv was imported.
- Summary statistics and data types were displayed to understand the initial structure and content of the data.

**2. Checking for Duplicates:**

- Identified and counted the number of duplicate rows in the dataset to avoid redundant information.
- Result: num_duplicates duplicates were found.

**3. Counting Rows with Zero MSRP:**

- Counted the number of rows where the Base.MSRP (Manufacturer's Suggested Retail Price) is zero to identify incomplete pricing data.
- Result: count_zero_msrp rows with zero MSRP were identified.

**4. Checking for Null Values:**

- Calculated the total number of null (NA) values in the dataset to address missing data.
- Result: total_na null values were found.

**5. Removing Null Values and Identifying Outliers:**

- Filtered out rows with null values to ensure data completeness.
- Identified the maximum Base.MSRP for each electric vehicle type to spot potential outliers.
- Result:The maximum MSRP for each vehicle type was summarized.

**6. Filtering Out Outliers and Zero MSRP:**

Removed the row with an extreme outlier Base.MSRP of 845,000 to prevent distortion in the analysis.
Also filtered out rows where Base.MSRP is zero to ensure only valid pricing data is considered.
By performing these steps, the dataset was cleaned and prepared, ensuring it is more accurate and reliable for subsequent analysis.

**Descriptive Statistics:**

- Compute summary statistics (mean, median, mode, standard deviation) for numerical variables.
- Generate frequency tables for categorical variables.

**Boxplot - Price Distribution vs. Electric Vehicle Type:**

This boxplot compares the price distribution of Battery Electric Vehicles (BEV) and Plug-in Hybrid Electric Vehicles (PHEV).
It shows that BEVs have a wider price range with a higher median price compared to PHEVs, which have a more compact price distribution with some outliers.

**Boxplot - Price Distribution vs. Electric Vehicle Type (Clean Alternative Fuel Vehicle Eligible):**

This boxplot illustrates the price distribution based on eligibility for Clean Alternative Fuel Vehicle incentives.
Vehicles eligible for the incentive tend to have a higher price range compared to those that are not eligible due to low battery range.

**Boxplot - Price Distribution vs. Electric Vehicle Type (Drive Range):**

This boxplot shows the distribution of drive range for clean alternative fuel vehicle eligible and non-eligible vehicles.
Eligible vehicles have a wider and higher range compared to non-eligible vehicles, which have a limited drive range.

**Scatterplot - Electric Vehicle Range Growth over Time:**

This scatterplot depicts the growth in electric vehicle range over time.
There is a clear trend showing that the electric range of vehicles has generally increased over the years.

**Boxplot - Price Distribution by Brand:**

This boxplot compares the price distribution of electric vehicles by different brands.
Tesla has the highest price range, followed by Porsche. Brands like KIA and Chrysler have lower price distributions.

**Scatterplot - Price vs. Range Correlation:**

This scatterplot shows the correlation between the vehicle price and its electric range.
There is a positive correlation where higher-priced vehicles tend to have a longer electric range.
These plots provide insights into the price and range variations of electric vehicles based on type, eligibility for incentives, brand, and over time.

**Data Visualization:**

- Create bar charts, histograms, and pie charts to visualize distributions.
- Use scatter plots and box plots to explore relationships between variables.
- Generate geographic maps to visualize EV distribution across the state.

**Comparative Analysis:**

- Compare BEVs and PHEVs in terms of electric range and base MSRP.
- Analyze differences in CAFV eligibility across makes and models.

**Correlation Analysis:**

- Calculate correlation coefficients for numerical variables.
- Use cross-tabulations to explore relationships between categorical and numerical variables.

**Geospatial Analysis (if applicable):**

- Use geographic mapping tools to identify clusters or patterns in vehicle registrations.

**Exploratory Data Analysis (EDA):**

-Perform in-depth analysis to identify trends, patterns, or outliers.
-Use advanced visualization techniques for deeper insights.

**Conclusion:**

By analyzing the electric vehicle registration data for Washington State, this project aims to uncover significant patterns and trends that can inform policy decisions and infrastructure planning. The insights gained from this analysis will contribute to the broader understanding of EV adoption and support the transition towards a more sustainable transportation future.

{{< pagebreak >}}

### 2024-06-21 code/processing-code/pre_process_tax_data.py
add models & convert explatoratory analysis to Python

@@ -1,24 +1,124 @@
-from pandas import read_csv
+from pandas import read_csv, set_option, merge
+from tqdm import tqdm # unnecessary dependency, adding as a meme 
+from numpy import mean
+from pyspark.sql import SparkSession
+set_option('display.max_columns', None)
 
-def convert_csv_to_parquet(year: int, dataset: str):
-    try:
-        print('Converting csv to parquet.')
-        csv_df = read_csv(f'{dataset}{year}.csv')
-        csv_df.to_parquet(f'{dataset}{year}.parquet')
-        print('Conversion successful!')
-    except Exception as e:
-        print('CSV to parquet conversion failed :', e)
+def data_dict_pre_process(filename: str):
+    strip_variable_whitespace_list = []
+    data_dict_df = read_csv(f'data/raw-data/{filename}.csv')
+    for idx, row in data_dict_df.iterrows():
+        variable = row['VARIABLE NAME'].strip()
+        strip_variable_whitespace_list.append(variable)
+    data_dict_df['VARIABLE NAME'] = strip_variable_whitespace_list
 
-years = [
-    '2011','2012','2013',
-    '2014','2015','2016',
-    '2017','2018','2019',
-    '2020','2021',]
+    data_dict_df_variable_names = data_dict_df['VARIABLE NAME'].tolist()
+    
+    tax_qualitative_variables = data_dict_df_variable_names[:19]
+    format_tax_qualitative_variables = []
+    for code in tqdm(tax_qualitative_variables):
+        format_tax_qualitative_variables.append(code)
+    
+    tax_numerical_variables = data_dict_df_variable_names[19:]
+    total_count_return_variables = []
+    total_amount_return_variables = []
+    for code in tqdm(tax_numerical_variables):
+        if code.startswith('N'):
+            total_count_return_variables.append(code)
+        else:
+            total_amount_return_variables.append(code)
 
-datasets_to_convert = [
-    'tax_zip_code_',
-    'data_dict',]
+    amount_return_variables_new = format_tax_qualitative_variables + total_amount_return_variables
+    count_return_variables_new = tax_qualitative_variables + total_count_return_variables
[…]

### 2024-06-22 code/eda-code/eda-ev.qmd
minor edits & begin working with geojson

---
title: "EV Exploratory Data Analysis"
date: today
output: html_document
execute: 
  error: False
  warning: False
---

# Setup

```{python}
#load needed packages. make sure they are installed.
from pandas import read_csv, set_option, merge, DataFrame, concat, read_json
import numpy as np
import matplotlib.pyplot as plt
import seaborn as sns
from sklearn.model_selection import train_test_split
from sklearn.linear_model import LinearRegression, LogisticRegression
from sklearn.preprocessing import LabelEncoder, StandardScaler
from sklearn.metrics import accuracy_score
import statsmodels.api as sm
import statsmodels
from statsmodels.graphics.gofplots import ProbPlot
from scipy import stats
import geojson
import folium
import json
set_option('display.max_columns', None)
```

Load the data.

```{python}
base_dir = '../../'
df = read_csv(base_dir + 'data/processed-data/brand_model_year_tax_data.csv')
columns_list = df.columns.tolist()
for i in columns_list:
    print(i.strip())
numeric_cols = columns_list[28:]
df.columns.tolist()
```

Vehicle Brand Aggregation
```{python}
sns.set_style('darkgrid')
zip_code_count = df.groupby(['ZIPCODE']).agg({
    'Vehicle Count': 'sum',
}).reset_index()
fig, axes = plt.subplots(2, 1, figsize=(15,15))
sns.histplot(ax=axes[0], data=zip_code_count, y=zip_code_count['Vehicle Count'])
log_zip_code_count = zip_code_count.copy()
log_zip_code_count['Vehicle Count'] = np.log(zip_code_count['Vehicle Count'])
sns.histplot(ax=axes[1], data=log_zip_code_count, y=log_zip_code_count['Vehicle Count'])
zip_code_count
```

```{python}
us_zips_df = read_csv(base_dir + 'data/raw-data/uszips.csv')
wa_zips = us_zips_df.loc[us_zips_df['state_id'] == 'WA']
wa_zips = wa_zips[['ZIPCODE', 'lat', 'lng', 'population']]
merge_zips = merge(wa_zips, zip_code_count, on='ZIPCODE')
# print(merge_zips)
test = base_dir + 'data/raw-data/counties.geojson'
with open(test) as f:
    gj = geojson.load(f)
    # print(gj['features'])
# countries = ["53"]
# new_list = []
# features = gj['features'][0]['properties']['STATEFP']
features = gj['features']
print(features[0])
# for i in features:
#     # print(i)
#     if i['properties']['STATEFP'] == '53':
#         new_list.append(i)
#     else:
#         pass
# final = json.dumps(new_list, indent=2)
# print(gj)
mapit = folium.Map(location=[122.511983, 37.77113])

for i in features:
    if i['properties']['STATEFP'] == '53':
        print(i)
        folium.Choropleth(geo_data=i['geometry']['coordinates'][0][0],
            data=merge_zips,
            columns=['lat', 'lng', 'Vehicle Count'],
            key_on='geometry.coordinates',
            fill_color='magma',
            line_color='magma'
        ).add_to(mapit)
# mapit.save('test.html')
# coordinates = []
# for zip_code in zip_code_count['ZIPCODE']:
#     print(zip_code)
#     geocode_result = geopandas.
    # geocode_result = gmaps.geocode(zip_code)
    # lat = geocode_result[0][‘geometry’][‘location’][‘lat’]
    # lng = geocode_result[0][‘geometry’][‘location’][‘lng’]
    # coordinates.append((lat, lng))
```

```{python} fig, axes = plt.subplots(2, 1, figsize=(15,15))
sns.boxplot(ax=axes[0], data=zip_code_count, x=zip_code_count['Vehicle Count'])
log_zip_code_count = zip_code_count.copy()
log_zip_code_count['Vehicle Count'] = np.log(zip_code_count['Vehicle Count'])
sns.boxplot(ax=axes[1], data=log_zip_code_count, x=log_zip_code_count['Vehicle Count'])
```

MAKE AND ELECTRIC VEHICLE TYPE
```{python}
sns.set_style('darkgrid')
make_ev_type = df.groupby(['Make']).agg({
    'Vehicle Count': 'sum',
}).reset_index()
fig, axes = plt.subplots(2, 1, figsize=(15,15))
sns.histplot(ax=axes[0], data=make_ev_type, y=make_ev_type['Vehicle Count'])
log_make_ev_type = make_ev_type.copy()
log_make_ev_type['Vehicle Count'] = np.log(make_ev_type['Vehicle Count'])
sns.histplot(ax=axes[1], data=log_make_ev_type, y=log_make_ev_type['Vehicle Count'])
```

```{python}
fig, axes = plt.subplots(2, 1, figsize=(15,15))
sns.boxplot(ax=axes[0], data=make_ev_type)
log_make_ev_type = make_ev_type.copy()
log_make_ev_type['Vehicle Count'] = np.log(make_ev_type['Vehicle Count'])
sns.boxplot(ax=axes[1], data=log_make_ev_type)
```

MAKE AND Clean Alt Fuel
```{python}
sns.set_style('darkgrid')
make_ev_type = df.groupby(['Make', 'Clean Alternative Fuel Vehicle (CAFV) Eligibility']).agg({
    'Vehicle Count': 'sum',
}).reset_index()
fig, axes = plt.subplots(2, 1, figsize=(15,15))
sns.histplot(ax=axes[0], data=make_ev_type, y=make_ev_type['Vehicle Count'])
log_make_ev_type = make_ev_type.copy()
log_make_ev_type['Vehicle Count'] = np.log(make_ev_type['Vehicle Count'])
sns.histplot(ax=axes[1], data=log_make_ev_type, y=log_make_ev_type['Vehicle Count'])
```

```{python}
fig, axes = plt.subplots(2, 1, figsize=(15,15))
sns.boxplot(ax=axes[0], data=make_ev_type, x=make_ev_type['Clean Alternative Fuel Vehicle (CAFV) Eligibility'], y=make_ev_type['Vehicle Count'])
log_make_ev_type = make_ev_type.copy()
log_make_ev_type['Vehicle Count'] = np.log(make_ev_type['Vehicle Count'])
sns.boxplot(ax=axes[1], data=log_make_ev_type, x=make_ev_type['Clean Alternative Fuel Vehicle (CAFV) Eligibility'], y=log_make_ev_type['Vehicle Count'])
```

Mean 
```{python}
brand_df = df[[
 'Residential energy tax credit amount', 'Total income amount', 'Vehicle Count'
]]

fig, axes = plt.subplots(1, 1, figsize=(15,15))
sns.heatmap(data=brand_df)
brand_df.corr()
# brand_df['Electric Range'] = np.log(brand_df['Electric Range'])
# sns.boxplot(ax=axes[1], data=brand_df, y=brand_df['Electric Range'])
```

```{python}
new_brand_df  = df.groupby(['ZIPCODE']).agg({
    'Total income amount': 'sum',
    'Electric Range': 'count'
}).reset_index(drop=True).rename(columns={
    'Electric Range': 'Vehicle Count'
})
fig, axes = plt.subplots(2, 1, figsize=(15,15))
new_brand_df['Total income amount'] = new_brand_df['Total income amount'].astype(int)
slope, intercept, r_value, p_value, std_err = stats.linregress(new_brand_df['Total income amount'],new_brand_df['Vehicle Count'])
sns.regplot(ax=axes[0], data=new_brand_df, y='Vehicle Count', x='Total income amount', 
    ci=None, 
    label=f"y={0:.1f}x+{1:.1f}\nr-sq: {r_value.round(3)}".format(slope, intercept)).legend(loc="best")
new_brand_df['Vehicle Count'] = np.log(new_brand_df['Vehicle Count'])
new_brand_df['Total income amount'] = np.log(new_brand_df['Total income amount'])
slope, intercept, r_value, p_value, std_err = stats.linregress(new_brand_df['Total income amount'],new_brand_df['Vehicle Count'])
sns.regplot(ax=axes[1],
    data=new_brand_df,
    x='Total income amount',
    y='Vehicle Count', 
    ci=None, 
    label=f"y={0:.1f}x+{1:.1f}\nr-sq: {r_value.round(3)}".format(slope, intercept)).legend(loc="best")

```

```{python}
base_dir = '../../'
df = read_csv(base_dir + 'data/processed-data/brand_model_year_tax_data.csv')
columns_list = df.columns.tolist()
for i in columns_list:
    print(i.strip())
# numeric_cols = columns_list[28:]
# df.columns.tolist()
```

ZIP CODE ONLY
```{python}
amount_brand_df = df.groupby(['ZIPCODE'])[numeric_cols].mean(numeric_only=True).reset_index()
brand_df = df.groupby(['ZIPCODE']).agg({
    'Vehicle Count': 'sum',
    'Electric Range': 'mean'
}).reset_index()
merged_df = merge(brand_df, amount_brand_df, on=['ZIPCODE'])
merged_df = merged_df.sort_values('Vehicle Count', ascending=False)

X_df = merged_df
Y_df = merged_df[['Vehicle Count']]

Y_df = np.log(Y_df)
X_df = merged_df.drop(['Vehicle Count', 'Electric Range', 'ZIPCODE'], axis=1)

X = np.array(X_df)
x = sm.add_constant(X)
Y = np.array(Y_df)
X_train, X_test, y_train, y_test = train_test_split(x,Y, 
                                   random_state=104,  
                                   test_size=0.2,
                                   shuffle=True) 
print(X_train.shape)
print(y_train.shape)
model = sm.OLS(y_train, X_train)
results = model.fit()
results.summary()

p_values = results.pvalues
p_df = DataFrame()
p_df['Feature'] = X_df.columns
p_df['p-value'] = results.pvalues[1:]
feat_list = []
p_val_list = []
for idx, row in p_df.iterrows():
    if row['p-value'] < .05:
        feat_list.append(row['Feature'])
        p_val_list.append(row['p-value'])
    else:
        continue
new_p_df = DataFrame()
new_p_df['Feature'] = feat_list
new_p_df['p-value'] = p_val_list
X = np.array(X_df[new_p_df['Feature']])
Y = np.array(Y_df)
x = sm.add_constant(X)

X_train, X_test, y_train, y_test = train_test_split(x,Y, 
                                   random_state=104,  
                                   test_size=0.2,
                                   shuffle=True) 
model2 = sm.OLS(y_train, X_train)
results2 = model2.fit()
results2.summary()

p3_values = results2.pvalues
new3 = DataFrame()
new3['Feature'] = new_p_df['Feature']
new3['p-value'] = p3_values[1:]
new3
feat_list3 = []
p_val_list3 = []
for idx, row in new3.iterrows():
    if row['p-value'] < .05:
        feat_list3.append(row['Feature'])
        p_val_list3.append(row['p-value'])
    else:
        continue
final3_df = DataFrame()
final3_df['Feature'] = feat_list3
final3_df['p-value'] = p_val_list3
final3_df
X = np.array(X_df[final3_df['Feature']])
Y = np.array(Y_df)
x = sm.add_constant(X)

X_train, X_test, y_train, y_test = train_test_split(x,Y, 
                                   random_state=104,  
                                   test_size=0.2,
                                   shuffle=True)

model3 = sm.OLS(y_train, X_train)
results3 = model3.fit()
Y_pred = results3.predict(X_train).astype(int)

features = ['constant'] + final3_df['Feature'].tolist()
X = DataFrame(X_train, columns=features)
Y = DataFrame(y_train, columns=['Vehicle Count'])
dataframe = concat([X, Y], axis=1)
dataframe
results3.summary()
features
```

```{python}
# model values
model_fitted_y = results3.fittedvalues

# model residuals
model_residuals = results3.resid

# normalized residuals
model_norm_residuals = results3.get_influence().resid_studentized_internal

# absolute squared normalized residuals
model_norm_residuals_abs_sqrt = np.sqrt(np.abs(model_norm_residuals))

# absolute residuals
model_abs_resid = np.abs(model_residuals)

# leverage, from statsmodels internals
model_leverage = results3.get_influence().hat_matrix_diag

# cook's distance, from statsmodels internals
model_cooks = results3.get_influence().cooks_distance[0]

plot_lm_1 = plt.figure()
plot_lm_1.axes[0] = sns.residplot(x=model_fitted_y, y=dataframe.columns[-1], data=dataframe,
                          lowess=True,
                          scatter_kws={'alpha': 0.5},
                          line_kws={'color': 'red', 'lw': 1, 'alpha': 0.8})

plot_lm_1.axes[0].set_title('Residuals vs Fitted')
plot_lm_1.axes[0].set_xlabel('Fitted values')
plot_lm_1.axes[0].set_ylabel('Residuals')
df = DataFrame()
df['res'] = model_residuals
df['fitted'] = model_fitted_y
df['count'] = dataframe['Vehicle Count']
# df['ZIPCODE'] = dataframe['ZIPCODE']
maxr = np.max(model_residuals)
df.sort_values('fitted', ascending=False)
```

#### Normal Q-Q Plot
This plot shows if the residuals are normally distributed. A good normal QQ plot has all of the residuals lying on or close to the red line.

```{python}
QQ = ProbPlot(model_norm_residuals)
plot_lm_2 = QQ.qqplot(line='45', alpha=0.5, color='#4C72B0', lw=1)
plot_lm_2.axes[0].set_title('Normal Q-Q')
plot_lm_2.axes[0].set_xlabel('Theoretical Quantiles')
plot_lm_2.axes[0].set_ylabel('Standardized Residuals')

# annotations
abs_norm_resid = np.flip(np.argsort(np.abs(model_norm_residuals)), 0)
abs_norm_resid_top_3 = abs_norm_resid[:3]
for r, i in enumerate(abs_norm_resid_top_3):
    plot_lm_2.axes[0].annotate(i,
                               xy=(np.flip(QQ.theoretical_quantiles, 0)[r],
                                   model_norm_residuals[i]))
```

#### Scale-Location
This plot is a way to check if the residuals suffer from non-constant variance, aka heteroscedasticity.

```{python}
plot_lm_3 = plt.figure()
plt.scatter(model_fitted_y, model_norm_residuals_abs_sqrt, alpha=0.5)
sns.regplot(
    x=model_fitted_y, 
    y=model_norm_residuals_abs_sqrt,
    scatter=False,
    ci=False,
    lowess=True,
    line_kws={'color': 'red', 'lw': 1, 'alpha': 0.8})
plot_lm_3.axes[0].set_title('Scale-Location')
plot_lm_3.axes[0].set_xlabel('Fitted values')
plot_lm_3.axes[0].set_ylabel('$\sqrt{|Standardized Residuals|}$')

# annotations
abs_sq_norm_resid = np.flip(np.argsort(model_norm_residuals_abs_sqrt), 0)
abs_sq_norm_resid_top_3 = abs_sq_norm_resid[:3]
for r, i in enumerate(abs_sq_norm_resid_top_3):
    plot_lm_3.axes[0].annotate(
        i,
        xy=(model_fitted_y[i],
        model_norm_residuals_abs_sqrt[i]))
```

#### Residuals vs. Leverage
Unlike outliers, which have an unusually large y value, leverage points have extreme x values. This may not seem so bad at face value, but it can have damaging effects on the model because the β coefficients are very sensitive to leverage points. The purpose of the Residuals vs Leverage plot is to identify these problematic observations.

```{python}
plot_lm_4 = plt.figure()
plt.scatter(model_leverage, model_norm_residuals, alpha=0.5)
sns.regplot(x=model_leverage, y=model_norm_residuals,
            scatter=False,
            ci=False,
            lowess=True,
            line_kws={'color': 'red', 'lw': 1, 'alpha': 0.8})
plot_lm_4.axes[0].set_xlim(0, np.max(model_leverage)+0.01)
plot_lm_4.axes[0].set_ylim(-3, 5)
plot_lm_4.axes[0].set_title('Residuals vs Leverage')
plot_lm_4.axes[0].set_xlabel('Leverage')
plot_lm_4.axes[0].set_ylabel('Standardized Residuals')

# annotations
leverage_top_3 = np.flip(np.argsort(model_cooks), 0)[:3]
for i in leverage_top_3:
    plot_lm_4.axes[0].annotate(i,
                                xy=(model_leverage[i],
                                    model_norm_residuals[i]))
```

```{python}
def cookdplot(model, ax=None):
    if ax is None:
        fig, ax = plt.subplots()

    cooks_d = model.get_influence().cooks_distance[0]

    ax.vlines(range(len(cooks_d)),0,cooks_d)

    # annotations
    cookd_top_3 = np.flip(np.argsort(cooks_d), 0)[:3]
    for i in cookd_top_3:
        ax.annotate(i, xy=(i, cooks_d[i]),color = 'C3')

    ax.set_title("Cook's Distance" , fontweight="bold")
    ax.set_xlabel('Obs. Number')
    ax.set_ylabel("Cook's distance")
    return ax

cookdplot(results3)
plt.show()
```

### SECOND MODEL
```{python}
# merged_df = merged_df.loc[(merged_df['Vehicle Count'] != 1.94591)]
# print(merged_df.loc[merged_df['Vehicle Count'] < 5].sort_values('Vehicle Count', ascending=False).head(30))
X_df = merged_df.copy()
Y_df = merged_df[['Vehicle Count']]
Y_df = np.log(Y_df)

X_features = [
#  'Adjust gross income (AGI) [8]',
#  'Total income amount',
 'State and local income tax refunds amount',
#  'Taxable pensions and annuities amount',
 'Number of farm returns',
 'Total statutory adjustments amount',
 'Self-employed (Keogh) retirement plans amount',
 'Total standard deduction amount',
#  'Basic standard deduction amount',
#  'Additional standard deduction amount',
#  'Total medical and dental expense deduction amount',
 'Limited state and local taxes',
 'Home mortgage from personal seller amount',
 'Alternative minimum tax amount',
 'Child and dependent care credit amount',
#  'Net premium tax credit amount',
#  'Qualified sick and family leave credit for leave taken before April 1, 2021 amount',
#  'Credited to next year’s estimated tax amount'
 ]
X_df = X_df[X_features]
feat_list = []
log_list = []
for feat in X_features:
    feat_list.append(feat)
    log_list.append(np.log(X_df[feat]))

# X_features = ['ZIPCODE',
#  'Adjust gross income (AGI) [8]',
#  'Total income amount',
#  'Total statutory adjustments amount',
#  'Home mortgage from personal seller amount',
#  'Alternative minimum tax amount']

X = X_df[X_features]

Y = Y_df
x = sm.add_constant(X)

X_train, X_test, y_train, y_test = train_test_split(x,Y, 
                                   random_state=104,  
                                   test_size=0.2,
                                   shuffle=True)

print(X_train)
model = sm.OLS(y_train, X_train)
results2mod = model.fit()
# Y_pred = results2mod.predict(X_test).astype(int)
results2mod.summary()
model1 = LinearRegression().fit(X_train, y_train)
model1.score(X_train, y_train)
# model1.score(X_test, y_test)
# results2mod.summary()

X = DataFrame(X_train, columns=['constant'] + X_features)
Y = DataFrame(y_train, columns=['Vehicle Count'])
dataframe = concat([X, Y], axis=1)
# dataframe
results2mod.summary()
# model1 = LinearRegression().fit(X_train, y_train)
# model1.score(X_train, y_train)
# model1.score(X_test, y_test)
# model.coef_
# model.intercept_
# Y_pred = model1.predict(X_test)
# plt.scatter(x=y_test, y=Y_pred)
# plt.plot(y_test, Y_pred, color='red')    
# sns.regplot(x=y_test, y=Y_pred)
```

```{python}
_# model values
model_fitted_y = results2mod.fittedvalues

# model residuals
model_residuals = results2mod.resid

# normalized residuals
model_norm_residuals = results2mod.get_influence().resid_studentized_internal

# absolute squared normalized residuals
model_norm_residuals_abs_sqrt = np.sqrt(np.abs(model_norm_residuals))

# absolute residuals
model_abs_resid = np.abs(model_residuals)

# leverage, from statsmodels internals
model_leverage = results2mod.get_influence().hat_matrix_diag
print(model_leverage)

# cook's distance, from statsmodels internals
model_cooks = results2mod.get_influence().cooks_distance[0]

plot_lm_1 = plt.figure()
plot_lm_1.axes[0] = sns.residplot(x=model_fitted_y, y=dataframe.columns[-1], data=dataframe,
                          lowess=True,
                          scatter_kws={'alpha': 0.5},
                          line_kws={'color': 'red', 'lw': 1, 'alpha': 0.8})

plot_lm_1.axes[0].set_title('Residuals vs Fitted')
plot_lm_1.axes[0].set_xlabel('Fitted values')
plot_lm_1.axes[0].set_ylabel('Residuals')
df = DataFrame()
df['res'] = model_residuals
df['fitted'] = model_fitted_y
df['count'] = dataframe['Vehicle Count']
df['index'] = df.index
maxr = np.max(model_residuals)
df.sort_values('res', ascending=True)
```

#### Normal Q-Q Plot
This plot shows if the residuals are normally distributed. A good normal QQ plot has all of the residuals lying on or close to the red line.

```{python}
QQ = ProbPlot(model_norm_residuals)
plot_lm_2 = QQ.qqplot(line='45', alpha=0.5, color='#4C72B0', lw=1)
plot_lm_2.axes[0].set_title('Normal Q-Q')
plot_lm_2.axes[0].set_xlabel('Theoretical Quantiles')
plot_lm_2.axes[0].set_ylabel('Standardized Residuals')

# annotations
abs_norm_resid = np.flip(np.argsort(np.abs(model_norm_residuals)), 0)
abs_norm_resid_top_3 = abs_norm_resid[:3]
for r, i in enumerate(abs_norm_resid_top_3):
    plot_lm_2.axes[0].annotate(i,
                               xy=(np.flip(QQ.theoretical_quantiles, 0)[r],
                                   model_norm_residuals[i]))
```

#### Scale-Location
This plot is a way to check if the residuals suffer from non-constant variance, aka heteroscedasticity.

```{python}
plot_lm_3 = plt.figure()
plt.scatter(model_fitted_y, model_norm_residuals_abs_sqrt, alpha=0.5)
sns.regplot(
    x=model_fitted_y, 
    y=model_norm_residuals_abs_sqrt,
    scatter=False,
    ci=False,
    lowess=True,
    line_kws={'color': 'red', 'lw': 1, 'alpha': 0.8})
plot_lm_3.axes[0].set_title('Scale-Location')
plot_lm_3.axes[0].set_xlabel('Fitted values')
plot_lm_3.axes[0].set_ylabel('$\sqrt{|Standardized Residuals|}$')

# annotations
abs_sq_norm_resid = np.flip(np.argsort(model_norm_residuals_abs_sqrt), 0)
abs_sq_norm_resid_top_3 = abs_sq_norm_resid[:3]
for r, i in enumerate(abs_sq_norm_resid_top_3):
    plot_lm_3.axes[0].annotate(
        i,
        xy=(model_fitted_y[i],
        model_norm_residuals_abs_sqrt[i]))
```

#### Residuals vs. Leverage
Unlike outliers, which have an unusually large y value, leverage points have extreme x values. This may not seem so bad at face value, but it can have damaging effects on the model because the β coefficients are very sensitive to leverage points. The purpose of the Residuals vs Leverage plot is to identify these problematic observations.

```{python}
plot_lm_4 = plt.figure()
plt.scatter(model_leverage, model_norm_residuals, alpha=0.5)
sns.regplot(x=model_leverage, y=model_norm_residuals,
            scatter=False,
            ci=False,
            lowess=True,
            line_kws={'color': 'red', 'lw': 1, 'alpha': 0.8})
plot_lm_4.axes[0].set_xlim(0, np.max(model_leverage)+0.01)
plot_lm_4.axes[0].set_ylim(-3, 5)
plot_lm_4.axes[0].set_title('Residuals vs Leverage')
plot_lm_4.axes[0].set_xlabel('Leverage')
plot_lm_4.axes[0].set_ylabel('Standardized Residuals')

# annotations
leverage_top_3 = np.flip(np.argsort(model_cooks), 0)[:3]
for i in leverage_top_3:
    plot_lm_4.axes[0].annotate(i,
                                xy=(model_leverage[i],
                                    model_norm_residuals[i]))
```

```{python}
def cookdplot(model, ax=None):
    if ax is None:
        fig, ax = plt.subplots()

    cooks_d = model.get_influence().cooks_distance[0]

    ax.vlines(range(len(cooks_d)),0,cooks_d)

    # annotations
    cookd_top_3 = np.flip(np.argsort(cooks_d), 0)[:3]
    for i in cookd_top_3:
        ax.annotate(i, xy=(i, cooks_d[i]),color = 'C3')

    ax.set_title("Cook's Distance" , fontweight="bold")
    ax.set_xlabel('Obs. Number')
    ax.set_ylabel("Cook's distance")
    return ax

cookdplot(results2mod)
plt.show()
```

### 2024-06-22 products/manuscript/Manuscript.qmd
add updates to Manuscript

@@
 - Summary statistics and data types were displayed to understand the initial structure and content of the data.
+- UPDATE (6.17.2024) - Downloaded IRS tax data to join with the EV data by zip code
+  - Tax data provides many financial variables that we will include in our models (i.e. can we predict the amount of EVs per zip code based off tax data?)
+- Added a geojson file to draw zip codes on a map, will use for visualizations but still figuring out how to elegantly filter the json
 
@@
 
+**Statistical Model**
+- First iteration of models
+- Produced a few statistically significant tax metrics for predicting the number of EVs in a given county
+- Washington demographics datasets could further enhance our analysis
+
 **Conclusion:**

### 2024-07-13 code/eda-code/eda-ev.qmd
updated eda file

@@
 ---
+**Electric Vehicle Data Analysis**
+
+# Introduction:
+
+This document is an exploratory data analysis of electric vehicle data.
+The data was obtained from the IRS and contains information on electric vehicles, their owners, and the tax benefits they receive.
+The data was cleaned and processed in R and saved as a CSV file.
+This document will load the data, explore it, and fit a linear regression model to predict the number of electric vehicles in a given zip code.
 
@@
 Vehicle Brand Aggregation
+
 ```{python}
@@
 }).reset_index()
+```
+The figure contains two histograms showing the distribution of vehicle counts by ZIP code.
+The top plot displays the original distribution, with most ZIP codes having a lower vehicle count and a few having very high counts, resulting in a skewed distribution.
+The bottom plot shows the log-transformed distribution, which spreads out the data and makes it easier to visualize lower counts.
+Both plots use a horizontal bar layout for easy comparison.
+
+```{python}
 fig, axes = plt.subplots(2, 1, figsize=(15,15))
@@
 ```
+
+The image shows two box plots of vehicle counts by ZIP code.
+The top plot displays the original vehicle counts, revealing most counts below 200 with many outliers up to 1000.
+The bottom plot uses log-transformed vehicle counts, normalizing the data and reducing the impact of outliers.
+Both plots highlight the distribution, central tendency, and variability of vehicle counts across ZIP codes.
 
@@
 ```
+
+### Conclusion
+
+In this analysis, two models were compared: the initial linear regression model and a refined model with selected features.
+The refined model showed improved performance with higher R-squared and lower AIC values, indicating a better fit.
+The selection of features based on statistical significance helped in enhancing the model's predictive accuracy.
+Ultimately, the refined model is preferred for predicting the number of electric vehicles in a given zip code due to its better performance metrics and more robust feature selection process.

### 2024-07-13 products/manuscript/Manuscript.qmd
updated eda file

@@
 ```{r}
-data_location <- here::here("starter-analysis-exercise","code","final_project")
-data <- read.csv("Electric_Vehicle_Population_Data.csv")
+data_location <- here::here("data","raw-data","Electric_Vehicle_Population_Data.csv")
+data <- read.csv(data_location)
 str(data)

### 2024-07-13 code/eda-code/eda-ev.qmd
comment out unfinished Python code block so QMD file renders correctly

@@
 
+### TO DO
 ```{python}
-us_zips_df = read_csv(base_dir + 'data/raw-data/uszips.csv')
-wa_zips = us_zips_df.loc[us_zips_df['state_id'] == 'WA']
-wa_zips = wa_zips[['ZIPCODE', 'lat', 'lng', 'population']]
-merge_zips = merge(wa_zips, zip_code_count, on='ZIPCODE')
-# print(merge_zips)
-test = base_dir + 'data/raw-data/counties.geojson'
-with open(test) as f:
-    gj = geojson.load(f)
+# us_zips_df = read_csv(base_dir + 'data/raw-data/uszips.csv')
+# wa_zips = us_zips_df.loc[us_zips_df['state_id'] == 'WA']
+# wa_zips = wa_zips[['ZIPCODE', 'lat', 'lng', 'population']]
+# merge_zips = merge(wa_zips, zip_code_count, on='ZIPCODE')
+# # print(merge_zips)
+# test = base_dir + 'data/raw-data/counties.geojson'
+# with open(test) as f:
+#     gj = geojson.load(f)
     # print(gj['features'])
@@
 # features = gj['features'][0]['properties']['STATEFP']
-features = gj['features']
-print(features[0])
+# features = gj['features']
+# print(features[0])
 # for i in features:
@@
 # print(gj)
-mapit = folium.Map(location=[122.511983, 37.77113])
+# mapit = folium.Map(location=[122.511983, 37.77113])
 
-for i in features:
-    if i['properties']['STATEFP'] == '53':
-        print(i)
-        folium.Choropleth(geo_data=i['geometry']['coordinates'][0][0],
-            data=merge_zips,
-            columns=['lat', 'lng', 'Vehicle Count'],
-            key_on='geometry.coordinates',
-            fill_color='magma',
-            line_color='magma'
-        ).add_to(mapit)
+# for i in features:
+#     if i['properties']['STATEFP'] == '53':
+#         print(i)
+#         folium.Choropleth(geo_data=i['geometry']['coordinates'][0][0],
+#             data=merge_zips,
+#             columns=['lat', 'lng', 'Vehicle Count'],
+#             key_on='geometry.coordinates',
+#             fill_color='magma',
+#             line_color='magma'
+#         ).add_to(mapit)
 # mapit.save('test.html')

### 2024-07-26 products/manuscript/Manuscript.qmd
Updated Readme and manuscript file

@@
 
-The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. 
-You don't need to have exactly these sections, but the content covering those sections should be addressed.
-
-This uses MS Word as output format.
-[See here](https://quarto.org/docs/output-formats/ms-word.html) for more information.
-You can switch to other formats, like html or pdf.
-See [the Quarto documentation](https://quarto.org/) for other formats.
-
 ```{r, echo=FALSE, message=FALSE}
@@
 ```
+Data Sources for the work:
 
@@
 -   https://registry.opendata.aws/
--   https://www.kaggle.com/datasets
-
 -   https://www.kaggle.com/datasets/jainaru/electric-vehicle-population
@@
 
-The dataset obtained from Kaggle provides detailed information about Battery Electric Vehicles (BEVs) and Plug-in Hybrid Electric Vehicles (PHEVs) registered through the Washington State Department of Licensing (DOL).
-It comprises 181,458 observations and 17 variables, encompassing key details such as Vehicle Identification Number (VIN), county, city, model year, electric range, and more.
+Electric vehicle market penetration is an important topic of interest for a variety of government agencies and automotive companies in terms of planning future charging infrastructure and inventory management.
+This project aims to aid this process by creation of a data driven model that can predict the number of electric vehicles in a given zip code based on a variety of factors.
+The data used in this project was sourced from the Washington State Department of Licensing (DOL) and includes information on Battery Electric Vehicles (BEVs) and Plug-in Hybrid Electric Vehicles (PHEVs) registered in Washington State.
+The data set contains 181,458 observations and 17 variables, including Vehicle Identification Number (VIN), county, city, model year, electric range, and more.
+The project will involve data cleaning, exploratory data analysis, and statistical modeling to identify key factors influencing electric vehicle adoption and predict the number of electric vehicles in a given zip code.
+The results of this project will provide valuable insights for policymakers, urban planners, and automotive companies to support the transition to sustainable transportation.
 
@@
 
-**- (a) Briefly describe what the data is, how it was collected, where you will get (or got) it from.
-How many observations do you have, what was measured?
-Anything else important to report about the data?**
+The dataset obtained from Kaggle provides detailed information about Battery Electric Vehicles (BEVs) and Plug-in Hybrid Electric Vehicles (PHEVs) registered through the Washington State Department of Licensing (DOL).
+It comprises 181,458 observations and 17 variables, encompassing key details such as Vehicle Identification Number (VIN), county, city, model year, electric range, and more.This document is an exploratory data analysis of electric vehicle data.
 
-The dataset obtained from Kaggle provides detailed information about Battery Electric Vehicles (BEVs) and Plug-in Hybrid Electric Vehicles (PHEVs) registered through the Washington State Department of Licensing (DOL).
-It comprises 181,458 observations and 17 variables, encompassing key details such as Vehicle Identification Number (VIN), county, city, model year, electric range, and more.
+A second set of data was obtained from the IRS and contains information on electric vehicles, their owners, and the tax benefits they receive.
+The data was cleaned and processed in R and saved as a CSV file.
+This data was then combined with thwe data from kaggle to conduct the analysis of data and model creation that predicts number of electric vehicles in a given zip code based on various input varaibles.
 
-**- (b) At this stage you are not required to already have and show the data, but if you do, even better.
-Then add a few lines of code which load the data and using some of the commands you learned about, provide summary descriptions of the data.**
+# Data set Analysis 
 ```{r}
@@
 data <- read.csv(data_location)
-str(data)
+output <- capture.output(str(data))
+cat(output[1:3], sep="\n")
 head(data)

### 2024-07-26 products/manuscript/supplement/Supplementary-Material.qmd
Updated Readme and manuscript file

---
title: "Supplement to Example Manuscript Template for a Data Analysis Project"
subtitle: ""
author: Andreas Handel
date: "`r Sys.Date()`"
format:
  pdf:
    toc: false
    number-sections: true
    highlight-style: github
#bibliography: ../../../assets/dataanalysis-references.bib
csl: ../../../assets/vancouver.csl
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

This shows some materials that could go into a supplementary file. Often you want/need references here too. You can use the same reference bib file for this and the main text (as done here) or have separate bib files.

For illustrative purposes, I'm doing the supplement as pdf. For this to work, you need a (La)TeX system installed. It's easy. Just follow [these steps](https://quarto.org/docs/output-formats/pdf-basics.html).

Of course you would choose the format based on needs.

I'm also using a different style for the references here. (vancouver vs apa in the main manuscript). Usually one would have the formatting of the references the same in those two documents, but I want to illustrate how easy it is to switch reference formatting styles, you just need to get the right CSL file and specify it in the YAML header. We could also have a seperate reference bibtext (`.bib`) file, but here we are using the same.

# Overview

A quick overview of what readers can find in the supplement.

# Code and file information

Explain here what each code/file is and does, and in which order (if any) users need to run thing to reproduce everything.
Essentially, give a full set of instructions to re-generate everything.

{{< pagebreak >}}

# Additional Method Details

Often, the main manuscript only allows for an overview description of the methods. Use the supplement to describe all your methods, models and approaches in a lot of detail. Reference specific parts of your code as needed.

{{< pagebreak >}}

# Additional results

Show additional results here. Those can be some useful exploratory/descriptive figures or tables, or results from additional analyses that didn't make it into the main text.

## Example additional result

@tbl-resulttable1 shows an additional table summarizing a model fit.

```{r}
#| label: tbl-resulttable1
#| tbl-cap: "Another fit table."
#| echo: FALSE
resulttable1 = readRDS(here("results","tables","resulttable1.rds"))
knitr::kable(resulttable1)
```

@fig-result2 shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result2
#| fig-cap: "Height and weight."
#| echo: FALSE
knitr::include_graphics(here("results","figures","height-weight.png"))
```

{{< pagebreak >}}

# Discussion

Any additional discussion regarding the supplementary material/findings.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

{{< pagebreak >}}

# References

### 2024-07-26 products/manuscript/supplement/Supplementary-Material.tex
Updated Readme and manuscript file

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Supplement to Example Manuscript Template for a Data Analysis Project},
  pdfauthor={Andreas Handel},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Supplement to Example Manuscript Template for a Data Analysis
Project}
\author{Andreas Handel}
\date{2024-07-24}

\begin{document}
\maketitle

This shows some materials that could go into a supplementary file. Often
you want/need references here too. You can use the same reference bib
file for this and the main text (as done here) or have separate bib
files.

For illustrative purposes, I'm doing the supplement as pdf. For this to
work, you need a (La)TeX system installed. It's easy. Just follow
\href{https://quarto.org/docs/output-formats/pdf-basics.html}{these
steps}.

Of course you would choose the format based on needs.

I'm also using a different style for the references here. (vancouver vs
apa in the main manuscript). Usually one would have the formatting of
the references the same in those two documents, but I want to illustrate
how easy it is to switch reference formatting styles, you just need to
get the right CSL file and specify it in the YAML header. We could also
have a seperate reference bibtext (\texttt{.bib}) file, but here we are
using the same.

\section{Overview}\label{overview}

A quick overview of what readers can find in the supplement.

\section{Code and file information}\label{code-and-file-information}

Explain here what each code/file is and does, and in which order (if
any) users need to run thing to reproduce everything. Essentially, give
a full set of instructions to re-generate everything.

\newpage{}

\section{Additional Method Details}\label{additional-method-details}

Often, the main manuscript only allows for an overview description of
the methods. Use the supplement to describe all your methods, models and
approaches in a lot of detail. Reference specific parts of your code as
needed.

\newpage{}

\section{Additional results}\label{additional-results}

Show additional results here. Those can be some useful
exploratory/descriptive figures or tables, or results from additional
analyses that didn't make it into the main text.

\subsection{Example additional result}\label{example-additional-result}

Table~\ref{tbl-resulttable1} shows an additional table summarizing a
model fit.

\begin{longtable}[]{@{}lrrrr@{}}

\caption{\label{tbl-resulttable1}Another fit table.}

\tabularnewline

\toprule\noalign{}
term & estimate & std.error & statistic & p.value \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
(Intercept) & 149.6997661 & 19.7518528 & 7.5790240 & 0.0001285 \\
Weight & 0.2277371 & 0.2708841 & 0.8407177 & 0.4282860 \\

\end{longtable}

Figure~\ref{fig-result2} shows a scatterplot figure produced by one of
the R scripts.

\begin{figure}

\centering{

\includegraphics[width=7in,height=\textheight]{../../../results/figures/height-weight.png}

}

\caption{\label{fig-result2}Height and weight.}

\end{figure}%

\newpage{}

\section{Discussion}\label{discussion}

Any additional discussion regarding the supplementary material/findings.

These papers {[}@mckay2020; @mckay2020a{]} are good examples of papers
published using a fully reproducible setup similar to the one shown in
this template.

\newpage{}

\section{References}\label{references}

\end{document}

### 2024-08-06 code/analysis-code/statistical_analysis.qmd
add updated models, manuscript, and final report

---
title: "Statistical Analysis"
date: today
output: html_document
execute: 
  error: False
  warning: False
---

## Import Libraries & Data
```{python}
import pandas as pd
from sklearn.model_selection import train_test_split
import re
pd.set_option('display.max_rows', None)

# Load the dataset
data = pd.read_csv('../../data/processed-data/brand_model_year_tax_data.csv').drop(
    [
        'The State Federal Information Processing System (FIPS) code',
        'The State associated with the ZIP code',
        'Number of returns [3]',
        'City',
        'State',
        'Model Year',
        'Make',
        'Model',
        'Electric Vehicle Type',
        'Clean Alternative Fuel Vehicle (CAFV) Eligibility',
    ],
    axis = 1
)

# Remove non-letters from columns names
regex = re.compile(r"\[|\]|<", re.IGNORECASE)

data.columns = [regex.sub("", col) if any(x in str(col) for x in set(('[', ']', '<'))) else col for col in data.columns.values]

data.columns = [re.sub(r'[0-9]', '', col).strip() for col in data.columns]
data.columns = [re.sub(r'/', '_', col) for col in data.columns]

# Group by 'ZIPCODE' and sum the 'Vehicle Count'
grouped_data = data.groupby('ZIPCODE')['Vehicle Count'].sum().reset_index()

# Selecting columns L to CY (indexes 11 to 102)
features_to_use = data.columns[3:-1]
print(features_to_use)

# Ensuring correct selection of numeric columns
numeric_features_to_use = data[features_to_use].select_dtypes(include=['number']).columns

# Grouping these features by ZIPCODE and summing them
grouped_features = data.groupby('ZIPCODE')[numeric_features_to_use].sum().reset_index()

# Merging the summed Vehicle Count with the grouped features on ZIPCODE
merged_data = grouped_data.merge(grouped_features, on='ZIPCODE')

# Extracting the features (X) and target (y)
X_grouped = merged_data[numeric_features_to_use]
y_grouped = merged_data['Vehicle Count']

# Splitting the data into training and testing sets
X_train_grouped, X_test_grouped, y_train_grouped, y_test_grouped = train_test_split(
    X_grouped, y_grouped, test_size=0.2, random_state=42)

```

## Preprocess, Transform, and Build Models
```{python}
import numpy as np
import pandas as pd
from sklearn.impute import SimpleImputer
import matplotlib.pyplot as plt
import seaborn as sns
from sklearn.linear_model import LinearRegression
from sklearn.ensemble import StackingRegressor, HistGradientBoostingRegressor, RandomForestRegressor
from sklearn.tree import DecisionTreeRegressor
import xgboost as xgb
from sklearn.metrics import mean_squared_error, r2_score
import scipy.stats as stats

# Define a function to log transform all features
def log_transform_all_features(data):
    return np.log1p(data)

# Impute missing values with the median and then log transform all features
def preprocess_and_log_transform_data(X):
    imputer = SimpleImputer(strategy='median')
    X_imputed = imputer.fit_transform(X)
    X_imputed_df = pd.DataFrame(X_imputed, columns=X.columns)
    X_transformed = log_transform_all_features(X_imputed_df)
    return X_transformed

# Preprocess and log transform the training and test datasets
X_train_transformed = preprocess_and_log_transform_data(X_train_grouped)
X_test_transformed = preprocess_and_log_transform_data(X_test_grouped)

# Define the models
models = {
    'HistGradientBoosting_Regressor': HistGradientBoostingRegressor(),
    'Decision_Tree_Regressor': DecisionTreeRegressor(max_depth=5, random_state=42),
    'XGBoost_Regressor': xgb.XGBRegressor(objective='reg:squarederror', n_estimators=100, learning_rate=0.1, max_depth=5, random_state=42),
    'Random_Forest_Regressor': RandomForestRegressor(random_state=42),
}

# Function to plot Q-Q plot, residuals distribution, and predicted vs. actual
def plot_model_diagnostics(y_true, y_pred, model_name):
    residuals = y_true - y_pred

    plt.figure(figsize=(18, 10))

    # Residuals Distribution
    plt.subplot(1, 2, 1)
    sns.histplot(residuals, kde=True, bins=30)
    plt.title(f'Residuals Distribution for {model_name}')
    plt.xlabel('Residuals')
    plt.ylabel('Frequency')

    # Predicted vs Actual
    plt.subplot(1, 2, 2)
    plt.scatter(y_true, y_pred, alpha=0.5, edgecolor='k')
    plt.plot([y_true.min(), y_true.max()], [y_true.min(), y_true.max()], 'r--', lw=2)
    plt.title(f'Predicted vs. Actual for {model_name}')
    plt.xlabel('Actual Values')
    plt.ylabel('Predicted Values')
    plt.grid(True)

    plt.tight_layout()
    plt.savefig(f'../../results/output/model_results/{model_name}.png')
    plt.show()

# Train each model and plot diagnostics
for name, model in models.items():
    model.fit(X_train_transformed, y_train_grouped)
    predictions = model.predict(X_test_transformed)

    # Print model evaluation metrics
    rmse = mean_squared_error(y_test_grouped, predictions, squared=False)
    r2 = r2_score(y_test_grouped, predictions)
    print(f"{name} - RMSE: {rmse}, R-squared: {r2}")

    # Plot model diagnostics
    plot_model_diagnostics(y_test_grouped, predictions, name)
```

## Variable Importance - Top 10
```{python}
from sklearn.inspection import permutation_importance
import pandas as pd

# Train the models on the transformed training data
xgboost_model = xgb.XGBRegressor(objective='reg:squarederror', n_estimators=100, learning_rate=0.1, max_depth=5, random_state=42)
hist_gb_model = HistGradientBoostingRegressor()
dec_tree_model = DecisionTreeRegressor(max_depth=5, random_state=42)
rand_for_model = RandomForestRegressor(random_state=42)

xgboost_model.fit(X_train_transformed, y_train_grouped)
hist_gb_model.fit(X_train_transformed, y_train_grouped)
dec_tree_model.fit(X_train_transformed, y_train_grouped)
rand_for_model.fit(X_train_transformed, y_train_grouped)

# Get feature names from the transformed data
feature_names = X_train_transformed.columns

# Extract feature importances from XGBoost
dec_tree_importances = dec_tree_model.feature_importances_
dec_tree_feature_importance = pd.DataFrame({
    'Feature': feature_names,
    'Importance': dec_tree_importances
}).sort_values(by='Importance', ascending=False)

# Extract feature importances from XGBoost
xgboost_importances = xgboost_model.feature_importances_
xgboost_feature_importance = pd.DataFrame({
    'Feature': feature_names,
    'Importance': xgboost_importances
}).sort_values(by='Importance', ascending=False)

# Compute feature importances using permutation importance for HistGradientBoostingRegressor
hist_gb_importances = permutation_importance(hist_gb_model, X_test_transformed, y_test_grouped, n_repeats=10, random_state=42)
hist_gb_feature_importance = pd.DataFrame({
    'Feature': feature_names,
    'Importance': hist_gb_importances.importances_mean
}).sort_values(by='Importance', ascending=False)

# Extract feature importances from XGBoost
rand_for_importances = rand_for_model.feature_importances_
rand_for_feature_importance = pd.DataFrame({
    'Feature': feature_names,
    'Importance': rand_for_importances
}).sort_values(by='Importance', ascending=False)

# Display the top 10 important features from each model
print("Top 10 Important Features from XGBoost:")
print(xgboost_feature_importance.head(10))

print("\nTop 10 Important Features from HistGradientBoostingRegressor (Permutation Importance):")
print(hist_gb_feature_importance.head(10))

print("\nTop 10 Important Features from Decision Tree:")
print(dec_tree_feature_importance.head(10))

print("\nTop 10 Important Features from Random Forest:")
print(rand_for_feature_importance.head(10))

rand_for_feature_importance_list = rand_for_feature_importance['Feature'].head(10).tolist()
print(rand_for_feature_importance_list)

# Plotting function
def plot_feature_importance(importances, model_name):
    plt.figure(figsize=(15, 10))
    plt.barh(importances['Feature'], importances['Importance'], color='skyblue')
    plt.xlabel('Importance')
    plt.ylabel('Feature')
    plt.title(f'Feature Importance - {model_name}')
    plt.gca().invert_yaxis()  # Invert y-axis to have the most important feature at the top
    plt.grid(True)
    plt.tight_layout()
    plt.savefig(f'../../results/output/variable_importance/{model_name}.png')

# Plotting feature importance for XGBoost
plot_feature_importance(xgboost_feature_importance.head(10), 'XGBoost')

# Plotting feature importance for HistGradientBoostingRegressor (Permutation Importance)
plot_feature_importance(hist_gb_feature_importance.head(10), 'HistGradientBoostingRegressor (Permutation Importance)')

# Plotting feature importance for HistGradientBoostingRegressor (Permutation Importance)
plot_feature_importance(rand_for_feature_importance.head(10), 'Random Forest Regressor')

# Plotting feature importance for HistGradientBoostingRegressor (Permutation Importance)
plot_feature_importance(dec_tree_feature_importance.head(10), 'Decision Tree Regressor')
```

## Compare Models
```{python}

from sklearn.metrics import mean_squared_error, mean_absolute_error, r2_score
import pandas as pd

# Function to evaluate models and compare various metrics
def compare_model_metrics(models, X_train, y_train, X_test, y_test):
    results = []

    for name, model in models.items():
        # Train the model
        model.fit(X_train, y_train)
        # Predict on the test set
        predictions = model.predict(X_test)
        # Calculate metrics
        rmse = mean_squared_error(y_test, predictions, squared=False)
        r2 = r2_score(y_test, predictions)
        mae = mean_absolute_error(y_test, predictions)
        mse = mean_squared_error(y_test, predictions)
        # Store the results
        results.append({
            'Model': name,
            'RMSE': rmse,
            'R-squared': r2,
            'MAE': mae,
            'MSE': mse
        })

    # Convert the results to a DataFrame for easy comparison
    return pd.DataFrame(results)

# Models to Compare
models_to_compare = {
    'XGBoost_Regressor': xgboost_model,  # Assuming xgboost_model is already defined
    'Decision_Tree_Regressor': DecisionTreeRegressor(max_depth=5, random_state=42),
    'HistGradientBoosting_Regressor': hist_gb_model,  # Assuming hist_gb_model is already defined
    'Random_Forest_Regressor': RandomForestRegressor(n_estimators=100, random_state=42),
}

# Evaluate and compare the models
model_comparison_metrics_df = compare_model_metrics(models_to_compare, X_train_transformed, y_train_grouped, X_test_transformed, y_test_grouped)

# Display the comparison results
print(model_comparison_metrics_df)
model_comparison_metrics_df.to_csv('../../results/output/model_comparison.csv', index=False)

```

## Visualize Model Comparison
```{python}
import matplotlib.pyplot as plt

# Function to visualize the model performance differences
def plot_model_performance_comparison(df, metric):
    plt.figure(figsize=(10, 6))
    df_sorted = df.sort_values(by=metric, ascending=False)
    plt.barh(df_sorted['Model'], df_sorted[metric], color='skyblue')
    plt.xlabel(metric)
    plt.ylabel('Model')
    plt.title(f'Model Comparison by {metric}')
    plt.gca().invert_yaxis()  # Invert y-axis to have the best performing model on top
    plt.grid(True)
    plt.savefig(f'../../results/output/model_comparison/{metric}.png')
    plt.show()

# Visualize RMSE comparison
plot_model_performance_comparison(model_comparison_metrics_df, 'RMSE')

# Visualize R-squared comparison
plot_model_performance_comparison(model_comparison_metrics_df, 'R-squared')

# Visualize MAE comparison
plot_model_performance_comparison(model_comparison_metrics_df, 'MAE')

# Visualize MSE comparison
plot_model_performance_comparison(model_comparison_metrics_df, 'MSE')
```

## Create Coefficients DataFrame
```{python}
# Import necessary libraries
import pandas as pd
import numpy as np

# Function to extract model coefficients and intercepts
def get_model_coefficients_and_intercepts(model, model_name, feature_names):
    if hasattr(model, 'coef_'):
        # Linear models and similar that have a coef_ attribute
        coefficients = model.coef_
    elif hasattr(model, 'feature_importances_'):
        # Tree-based models with feature_importances_ attribute
        coefficients = model.feature_importances_
    else:
        coefficients = np.nan  # Placeholder for models without accessible coefficients

    coefficients_df = pd.DataFrame({
        'Feature': feature_names,
        'Coefficient': coefficients,
        'Model': model_name
    })

    return pd.concat([coefficients_df], ignore_index=True)

# Get feature names
feature_names = X_train_transformed.columns

# Initialize models (assuming they are already trained)
models = {
    'XGBoost_Regressor': xgboost_model,
    'Decision_Tree_Regressor': dec_tree_model,
    'HistGradientBoosting_Regressor': hist_gb_model,
    'Random_Forest_Regressor': rand_for_model,
}

# Collect coefficients for each model
coefficients_with_intercepts_list = []

for model_name, model in models.items():
    coefficients_with_intercepts_list.append(get_model_coefficients_and_intercepts(model, model_name, feature_names))

# Combine all coefficients into one DataFrame
coefficients_df = pd.concat(coefficients_with_intercepts_list, ignore_index=True)

# Display the coefficients DataFrame
coefficients_df = coefficients_df.dropna()
coefficients_df.to_csv('../../results/output/model_coefficients.csv', index=False)
```

### 2024-08-06 code/eda-code/eda-ev-dataset.qmd
add updated models, manuscript, and final report

--- 2024-06-16 code/eda-code/data_exploration_part_2.qmd
@@
 # Import Data, Check Descriptive Statistics & Data Types
-data_location <- here::here("starter-analysis-exercise","code","final_project","Electric_Vehicle_Population_Data.csv")
+data_location <- here::here("data","raw-data","Electric_Vehicle_Population_Data.csv")
 ev_data <- read.csv(data_location)

### 2024-08-06 code/eda-code/eda-ev-tax.qmd
add updated models, manuscript, and final report

---
title: "EV Exploratory Data Analysis"
date: today
output: html_document
execute: 
  error: False
  warning: False
---
**Electric Vehicle Data Analysis**

# Introduction:

This document is an exploratory data analysis of electric vehicle data. The data was obtained from the IRS and contains information on electric vehicles, their owners, and the tax benefits they receive. The data was cleaned and processed in R and saved as a CSV file. This document will load the data, explore it, and fit a linear regression model to predict the number of electric vehicles in a given zip code.

# Setup

```{python}
import pandas as pd
from sklearn.model_selection import train_test_split
import re

# Load the dataset
data = pd.read_csv('../../data/processed-data/brand_model_year_tax_data.csv').drop(
    [
        'The State Federal Information Processing System (FIPS) code',
        'The State associated with the ZIP code',
        'Number of returns [3]',
        'City',
        'State',
        'Model Year',
        'Make',
        'Model',
        'Electric Vehicle Type',
        'Clean Alternative Fuel Vehicle (CAFV) Eligibility',
    ],
    axis = 1
)

# Remove non-letters from columns names
regex = re.compile(r"\[|\]|<", re.IGNORECASE)

data.columns = [regex.sub("", col) if any(x in str(col) for x in set(('[', ']', '<'))) else col for col in data.columns.values]

data.columns = [re.sub(r'[0-9]', '', col).strip() for col in data.columns]
data.columns = [re.sub(r'/', '_', col) for col in data.columns]

# Group by 'ZIPCODE' and sum the 'Vehicle Count'
grouped_data = data.groupby('ZIPCODE')['Vehicle Count'].sum().reset_index()

# Selecting columns L to CY (indexes 11 to 102)
features_to_use = data.columns[2:]

# Ensuring correct selection of numeric columns
numeric_features_to_use = data[features_to_use].select_dtypes(include=['number']).columns

# Grouping these features by ZIPCODE and summing them
grouped_features = data.groupby('ZIPCODE')[numeric_features_to_use].sum().reset_index().drop('Vehicle Count', axis = 1)

# Merging the summed Vehicle Count with the grouped features on ZIPCODE
merged_data = grouped_data.merge(grouped_features, on='ZIPCODE')

# Extracting the features (X) and target (y)
X_grouped = merged_data[numeric_features_to_use]
y_grouped = merged_data['Vehicle Count']

# Splitting the data into training and testing sets
X_train_grouped, X_test_grouped, y_train_grouped, y_test_grouped = train_test_split(
    X_grouped, y_grouped, test_size=0.3, random_state=42)

```

```{python}
import numpy as np
import pandas as pd
from sklearn.impute import SimpleImputer
import matplotlib.pyplot as plt
import seaborn as sns

# Define a function to log transform all features
def log_transform_all_features(data):
    return np.log(data)

# Impute missing values with the median and then log transform all features
def preprocess_and_log_transform_data(X):
    imputer = SimpleImputer(strategy='median')
    X_imputed = imputer.fit_transform(X)
    X_imputed_df = pd.DataFrame(X_imputed, columns=X.columns)
    X_transformed = log_transform_all_features(X_imputed_df)
    return X_transformed

# Preprocess and log transform the training and test datasets
X_train_transformed = preprocess_and_log_transform_data(X_train_grouped)
X_test_transformed = preprocess_and_log_transform_data(X_test_grouped)

# Plot histogram for each feature in the transformed dataset
def plot_transformed_feature_histograms(data, data1):

    for feature in data.columns:
        fig, ax = plt.subplots(1, 2, figsize=(10,10))
        ax1 = plt.subplot(2,1,1)
        ax1 = sns.histplot(data[feature], kde=True, bins=50)
        ax2 = plt.subplot(2,1,2)
        ax2 = sns.histplot(data1[feature], kde=True, bins=50)
        ax1.title.set_text(f'Histogram of {feature}')
        ax2.title.set_text(f'Log-Transformed Histogram of {feature}')
        plt.xlabel(feature)
        plt.ylabel('Frequency')
        plt.grid(True)
        plt.savefig(f'../../results/figures/histograms/{feature}_histogram.png')

# Plot the histograms for the log-transformed features
plot_transformed_feature_histograms(X_train_grouped, X_train_transformed)
```

```{python}
# Plot boxplot for each feature in the transformed dataset
def plot_transformed_feature_boxplots(data, data1):

    for feature in data.columns:
        fig, ax = plt.subplots(1, 2, figsize=(10,10))
        ax1 = plt.subplot(2,1,1)
        ax1 = sns.boxplot(data[feature])
        ax2 = plt.subplot(2,1,2)
        ax2 = sns.boxplot(data1[feature])
        ax1.title.set_text(f'Boxplot of {feature}')
        ax2.title.set_text(f'Log-Transformed Boxplot of {feature}')
        plt.xlabel(feature)
        plt.ylabel('Frequency')
        plt.grid(True)
        plt.tight_layout()
        plt.savefig(f'../../results/figures/boxplots/{feature}_boxplot.png')

# Plot the boxplots for the log-transformed features
plot_transformed_feature_boxplots(X_train_grouped, X_train_transformed)
```

### 2024-08-06 code/processing-code/pre_process_tax_data.qmd
add updated models, manuscript, and final report

---
title: "Pre-Process Tax Data"
date: today
output: html_document
execute: 
  error: False
  warning: False
---

## Import Libraries
```{python}
from pandas import read_csv, set_option, merge
from numpy import mean
set_option('display.max_columns', None)
```

## Reformat Data Dictionary Names
```{python}
def data_dict_pre_process(filename: str):
    strip_variable_whitespace_list = []
    data_dict_df = read_csv(f'../../data/raw-data/{filename}.csv')
    for idx, row in data_dict_df.iterrows():
        variable = row['VARIABLE NAME'].strip()
        strip_variable_whitespace_list.append(variable)
    data_dict_df['VARIABLE NAME'] = strip_variable_whitespace_list

    data_dict_df_variable_names = data_dict_df['VARIABLE NAME'].tolist()

    tax_qualitative_variables = data_dict_df_variable_names[:19]
    format_tax_qualitative_variables = []
    for code in tax_qualitative_variables:
        format_tax_qualitative_variables.append(code)

    tax_numerical_variables = data_dict_df_variable_names[19:]
    total_count_return_variables = []
    total_amount_return_variables = []
    for code in tax_numerical_variables:
        if code.startswith('N'):
            total_count_return_variables.append(code)
        else:
            total_amount_return_variables.append(code)

    amount_return_variables_new = format_tax_qualitative_variables + total_amount_return_variables
    count_return_variables_new = tax_qualitative_variables + total_count_return_variables
    description_amount_data_dict_list = data_dict_df['DESCRIPTION'].loc[data_dict_df['VARIABLE NAME'].isin(amount_return_variables_new)].tolist()
    description_count_data_dict_list = data_dict_df['DESCRIPTION'].loc[data_dict_df['VARIABLE NAME'].isin(count_return_variables_new)].tolist()

    amount_data_dict_df = data_dict_df.loc[data_dict_df['VARIABLE NAME'].isin(amount_return_variables_new)]
    count_data_dict_df = data_dict_df.loc[data_dict_df['VARIABLE NAME'].isin(count_return_variables_new)]
    amount_data_dict_df.to_csv('../../data/intermediate-data/data_dict_variable_total_amount.csv', index=False)
    count_data_dict_df.to_csv('../../data/intermediate-data/data_dict_variable_total_number_returns.csv', index=False)
```

## Preprocess Tax Data
```{python}
def tax_data_pre_process(filename: str, state: str):
    tax_data_df = read_csv(f'../../data/raw-data/{filename}.csv').drop('agi_stub', axis=1)
    tax_data_df = tax_data_df.loc[tax_data_df['STATE'] == f'{state}']

    amount_data_dict_df = read_csv(f'../../data/intermediate-data/data_dict_variable_total_amount.csv')
    amount_data_dict_df_variables_list = amount_data_dict_df['VARIABLE NAME'].tolist()
    description_amount_list = amount_data_dict_df['DESCRIPTION'].tolist()
    total_amount_tax_data_df = tax_data_df.loc[:, tax_data_df.columns.isin(amount_data_dict_df_variables_list)]
    total_amount_tax_data_df.columns = description_amount_list
    total_amount_tax_data_df = total_amount_tax_data_df.rename(
        columns={
            '5-digit Zip code': 'ZIPCODE'
        }
    )
    total_amount_tax_data_df.to_csv(f'../../data/intermediate-data/{state}_total_amount_tax_data.csv', index=False)

    number_data_dict_df = read_csv(f'../../data/intermediate-data/data_dict_variable_total_number_returns.csv')
    number_data_dict_df_variables_list = number_data_dict_df['VARIABLE NAME'].tolist()
    description_number_list = number_data_dict_df['DESCRIPTION'].tolist()
    total_number_tax_data_df = tax_data_df.loc[:, tax_data_df.columns.isin(number_data_dict_df_variables_list)]
    total_number_tax_data_df.columns = description_number_list
    total_number_tax_data_df = total_number_tax_data_df.rename(
        columns={
            '5-digit Zip code': 'ZIPCODE'
        }
    )
    total_number_tax_data_df.to_csv(f'../../data/intermediate-data/{state}_total_number_returns_tax_data.csv', index=False)
```

## Preprocess EV Data
```{python}
def ev_data_pre_process(state: str):
    ev_data_df = read_csv('../../data/raw-data/Electric_Vehicle_Population_Data.csv').rename(columns={
        'Postal Code': 'ZIPCODE'
    })
    ev_data_df['ZIPCODE'] = ev_data_df['ZIPCODE'].fillna(0).astype(int)
    null_values = ev_data_df.isnull().sum()
    ev_data_df.dropna(inplace=True)
    ev_data_df_2020 = ev_data_df.loc[(ev_data_df['Model Year'] <= 2020) & (ev_data_df['Electric Range'] > 0) & (ev_data_df['Model Year'] > 2010) & (ev_data_df['State'] == f'{state}')]
    ev_duplicated_df = ev_data_df[ev_data_df.duplicated()]

    zip_code_count_brand_model_year = ev_data_df_2020.groupby(['City', 'State', 'ZIPCODE', 'Model Year', 'Make', 'Model', 'Electric Vehicle Type', 'Clean Alternative Fuel Vehicle (CAFV) Eligibility']).agg({
        'Electric Range': 'mean',
        'County': 'count',
    }).reset_index()
    zip_code_count_brand_model = ev_data_df_2020.groupby(['City', 'State', 'ZIPCODE', 'Make', 'Model', 'Electric Vehicle Type', 'Clean Alternative Fuel Vehicle (CAFV) Eligibility']).agg({
        'Electric Range': 'mean',
        'County': 'count',
    }).reset_index()
    zip_code_count_brand = ev_data_df_2020.groupby(['City', 'State', 'ZIPCODE', 'Make', 'Electric Vehicle Type', 'Clean Alternative Fuel Vehicle (CAFV) Eligibility']).agg({
        'Electric Range': 'mean',
        'County': 'count',
    }).reset_index()
    zip_code_count_brand_model_year.to_csv(f'../../data/intermediate-data/{state}_zip_code_brand_model_year.csv', index=False)
    zip_code_count_brand_model.to_csv(f'../../data/intermediate-data/{state}_zip_code_brand_model.csv', index=False)
    zip_code_count_brand.to_csv(f'../../data/intermediate-data/{state}_zip_code_brand.csv', index=False)
```

## Join Tax/EV Data
```{python}
def join_tax_ev_data(state: str):
    zip_code_count_brand_model_year = read_csv(f'../../data/intermediate-data/{state}_zip_code_brand_model_year.csv')
    zip_code_count_brand_model= read_csv(f'../../data/intermediate-data/{state}_zip_code_brand_model.csv')
    zip_code_count_brand = read_csv(f'../../data/intermediate-data/{state}_zip_code_brand.csv')
    amount_tax_data = read_csv(f'../../data/intermediate-data/{state}_total_amount_tax_data.csv')

    final_zip_code_count_brand_model_year = merge(zip_code_count_brand_model_year, amount_tax_data, on='ZIPCODE').rename(
        columns={
            'County': 'Vehicle Count'
        }
    )
    final_zip_code_count_brand_model = merge(zip_code_count_brand_model, amount_tax_data, on='ZIPCODE').rename(
        columns={
            'County': 'Vehicle Count'
        }
    )
    final_zip_code_count_brand = merge(zip_code_count_brand, amount_tax_data, on='ZIPCODE').rename(
        columns={
            'County': 'Vehicle Count'
        }
    )
    final_zip_code_count_brand_model_year = final_zip_code_count_brand_model_year.drop(
        [
            'Adjust gross income (AGI) [8]',
            'Number of volunteer prepared returns with Earned Income Credit [5]',
            'Total income amount',
            'Total statutory adjustments amount',
            'Total standard deduction amount',
            'Total itemized deductions amount',
            'Amount of AGI for itemized returns',
            'Total taxes paid amount',
            'Taxable income amount',
            'Income tax before credits amount',
            'Alternative minimum tax amount',
            'Total tax credits amount',
            'Total premium tax credit amount',
            'Total tax payments amount',
            'Income tax after credits amount   ',
            'Total tax liability amount [18]  ',
            'Tax due at time of filing amount [19]  ',
            'Total overpayments amount',
            'Overpayments refunded amount [20]  ',
        ],
        axis = 1
    )
    final_zip_code_count_brand_model_year.to_csv('../../data/processed-data/brand_model_year_tax_data.csv', index=False)
    final_zip_code_count_brand_model.to_csv('../../data/processed-data/brand_model_tax_data.csv', index=False)
    final_zip_code_count_brand.to_csv('../../data/processed-data/brand_tax_data.csv', index=False)
```

## Execute Functions For Final Dataset
```{python}
data_dict_pre_process(filename='data_dict2021')
tax_data_pre_process(filename='tax_data_2021', state='WA')
ev_data_pre_process(state='WA')
join_tax_ev_data(state='WA')
```

### 2024-08-06 products/manuscript/Manuscript.qmd
add updated models, manuscript, and final report

@@
 ---
-title: "Example Manuscript Template for a Data Analysis Project"
+title: "Using IRS Data to Predict Electric Vehicle Count"
 subtitle: ""
@@
 
-```{r, echo=FALSE, message=FALSE}
-# load a few R packages
-library(here)
-library(knitr)
-```
-Data Sources for the work:
-
--   https://www.google.com/publicdata/directory
--   https://www.data.gov/
--   https://github.com/awesomedata/awesome-public-datasets
--   https://registry.opendata.aws/
--   https://www.kaggle.com/datasets/jainaru/electric-vehicle-population
-
 # Summary/Abstract
-_Write a summary of your project._
 
@@
 # Data set Analysis 
-```{r}
-data_location <- here::here("data","raw-data","Electric_Vehicle_Population_Data.csv")
-data <- read.csv(data_location)
-output <- capture.output(str(data))
-cat(output[1:3], sep="\n")
-head(data)
-summary(data)
-colSums(is.na(data))
+```{python}
+import pandas as pd
+pd.set_option('display.max_rows', None)
+
+# Load the dataset
+data = pd.read_csv('../../data/processed-data/brand_model_year_tax_data.csv').drop(
+    [
+        'The State Federal Information Processing System (FIPS) code',
+        'The State associated with the ZIP code',
+        'Number of returns [3]'
+    ],
+    axis = 1
+)
+data.head(10)
 ```
@@
 
-**Geospatial Analysis (if applicable):**
-
-- Use geographic mapping tools to identify clusters or patterns in vehicle registrations.
-
 **Exploratory Data Analysis (EDA):**
@@
 - First iteration of models
-- Produced a few statistically significant tax metrics for predicting the number of EVs in a given county
+- Produced a few statistically significant tax metrics for predicting the number of EVs in a given zip code
 - Washington demographics datasets could further enhance our analysis

### 2024-08-06 products/manuscript/report.qmd
add updated models, manuscript, and final report

---
title: "Using IRS Data to Predict Electric Vehicle Count"
author: Sri Lakshmi Chundru, Collin Real, Joaquin Ramirez, Seth Harris
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/apa.csl
---

## Abstract
This study investigates the use of advanced machine learning models to predict vehicle count by zip code, utilizing a dataset enriched with socio-economic, demographic, and regional features. The primary aim is to accurately estimate vehicle counts, aiding in efficient resource distribution and urban planning. We implemented and evaluated several models, including Decision Tree, Random Forest, XGBoost, and HistGradientBoosting, focusing on their accuracy, interpretability, and computational efficiency.

Each model was trained and validated using cross-validation and a hold-out test set, with performance metrics such as Mean Absolute Error (MAE), Mean Squared Error (MSE), and R-squared used to assess their effectiveness. Among the models, the Random Forest model demonstrated superior performance, offering the best balance of predictive accuracy and model robustness. It effectively captured the complex interactions between the features and the target variable, outperforming the other models in terms of both accuracy and generalization.

The results highlight the Random Forest model's effectiveness in handling diverse and complex datasets, making it a valuable tool for predicting vehicle counts in different zip codes. This study provides critical insights into the practical applications of machine learning in urban planning and resource management, particularly in the context of transportation and infrastructure development.

## Introduction
As electric vehicles (EVs) become increasingly prevalent, understanding their distribution and associated factors at a granular level, such as by zip code, is crucial for infrastructure planning and policy development. In this study, we sought to predict vehicle counts, with a particular focus on EVs, across various zip codes using a comprehensive dataset combining electric vehicle registration data with IRS tax data. This integration provides a robust foundation for analyzing socio-economic and demographic influences on vehicle distribution.

The dataset was curated by merging electric vehicle registration records, which provide detailed counts of EVs by zip code, with IRS tax data, offering insights into the income levels and financial characteristics of the residents in each zip code. This integration was facilitated by matching records based on zip codes, creating a unified dataset that captures a wide array of features, including income levels, tax filings, and vehicle counts.

Before analysis, significant data cleaning was necessary to ensure the dataset's accuracy and consistency. The initial data contained various inconsistencies, particularly in column names and formats. We employed regular expressions, a powerful tool for text manipulation, to standardize column names, correcting typographical errors, and ensuring uniformity in naming conventions. This step was crucial for seamless data manipulation and analysis, allowing for accurate merging and comparison of data across different sources.

Furthermore, to focus our analysis on a specific geographic region and manage data complexity, we filtered the dataset to include only records from the state of Washington. This state was chosen due to its diverse socio-economic landscape and significant adoption of electric vehicles, providing a rich dataset for analysis. The filtering process involved selecting rows where the state code matched Washington's, ensuring that the analysis remained relevant to the regional context.

By combining these datasets and refining them through meticulous cleaning and filtering, we prepared a comprehensive and reliable dataset. This dataset serves as the foundation for training machine learning models aimed at predicting vehicle counts by zip code, with a particular emphasis on understanding the factors influencing the distribution of electric vehicles.

## Methods
To address skewness and non-normal distributions in our predictor variables, we applied a log transformation to each feature in the dataset. This transformation helps to stabilize variance, minimize the impact of outliers, and create a more normally distributed set of predictor variables. Specifically, we used the natural logarithm function, applying the transformation as log⁡(x+1)log(x+1) to handle zero values gracefully. This preprocessing step was crucial for enhancing the performance and interpretability of our machine learning models.

### Model Building

1. Decision Tree Regressor
The Decision Tree Regressor is a simple yet powerful model that splits the data based on feature values, recursively partitioning the data into subsets with similar outcomes. For this model, we set parameters such as the maximum depth of the tree and the minimum number of samples required to split a node. We performed hyperparameter tuning to find the optimal tree depth and the minimum number of samples per leaf, balancing model complexity and overfitting.

2. Random Forest Regressor
The Random Forest Regressor builds upon the Decision Tree model by constructing a multitude of decision trees during training and outputting the mean prediction of the individual trees. This ensemble method improves predictive accuracy and controls overfitting. We specified key parameters such as the number of trees in the forest (n_estimators), the maximum depth of each tree, and the maximum number of features considered for splitting at each node. We also used bootstrap sampling, allowing each tree to be trained on a random subset of the data, enhancing model robustness.

3. XGBoost Regressor
XGBoost (Extreme Gradient Boosting) is an advanced implementation of gradient boosting designed for speed and performance. It sequentially builds decision trees, where each subsequent tree aims to correct the errors of the previous ones. For XGBoost, we tuned hyperparameters such as the learning rate (eta), the maximum depth of trees (max_depth), the number of boosting rounds (n_estimators), and regularization parameters (lambda and alpha) to prevent overfitting. Additionally, we used subsampling of rows (subsample) and columns (colsample_bytree) to introduce variability in the model, further preventing overfitting.

4. HistGradientBoosting Regressor
The HistGradientBoosting Regressor, part of the scikit-learn library, is a fast and scalable implementation of gradient boosting using histogram-based binning. This method is particularly efficient for large datasets. We configured parameters including the maximum number of iterations (max_iter), learning rate, maximum depth of the trees, and the number of bins (max_bins) for histogram binning. The model also included support for early stopping, which halts training when the validation loss stops improving, thus preventing overfitting and reducing training time.

## Results
This section presents the comparative analysis of various machine learning models used to predict vehicle count by zip code. The models evaluated include XGBoost, Decision Tree, HistGradientBoosting, and Random Forest. The primary evaluation metrics considered were Root Mean Squared Error (RMSE) and R-squared, which provide insights into the accuracy and explanatory power of the models.

### Model Performance Comparison
#### XGBoost
  - RMSE: 71.27 <br/>
  - R-squared: 0.91 <br/>
  - XGBoost demonstrated strong predictive capabilities, achieving a relatively low RMSE of 71.27. The model's R-squared value of 0.91 indicates that it explains 91% of the variance in the vehicle count data, suggesting good model fit and predictive accuracy. <br/>
![XGBoost](../../results/output/model_results/XGBoost_Regressor.png)

#### Decision Tree
  - RMSE: 100.77<br/>
  - R-squared: 0.83<br/>
  -  The Decision Tree model had the highest RMSE among the evaluated models at 100.77, reflecting less precise predictions. Its R-squared value of 0.83 indicates that it explains 83% of the variance, making it the least accurate model in this comparison.
![Decision Tree](../../results/output/model_results/Decision_Tree_Regressor.png)

#### HistGradientBoosting
  -  RMSE: 83.42<br/>
  -  R-squared: 0.88<br/>
  -  The HistGradientBoosting model offered a moderate performance with an RMSE of 83.42 and an R-squared value of 0.88. While better than the Decision Tree model, it still lagged behind XGBoost and Random Forest in terms of prediction accuracy and variance explanation.
![HistGradientBoosting](../../results/output/model_results/HistGradientBoosting_Regressor.png)

#### Random Forest
  -  RMSE: 62.16<br/>
  -  R-squared: 0.935<br/>
  -  The Random Forest model outperformed all other models, achieving the lowest RMSE of 62.16 and the highest R-squared value of 0.935. These metrics indicate that the Random Forest model not only provided the most accurate predictions but also explained 93.5% of the variance in vehicle counts, making it the most reliable model in this analysis.
![Random Forest](../../results/output/model_results/Random_Forest_Regressor.png)

### Best Model Selection

Based on the comparative analysis of the models, the Random Forest model emerged as the best performing model. It achieved the lowest RMSE and the highest R-squared, demonstrating superior accuracy and generalization capability compared to the other models. The Random Forest model's ensemble nature allows it to effectively capture complex patterns in the data, reduce overfitting through bootstrapping, and handle interactions between predictor variables more adeptly. These attributes collectively contributed to its superior performance, making it the most suitable choice for predicting vehicle counts by zip code in this study.

The excellent performance of the Random Forest model suggests it is well-suited for applications requiring high prediction accuracy and robustness, particularly in urban planning and resource allocation scenarios where accurate vehicle count predictions are crucial.

## Discussion
The objective of this study was to develop machine learning models capable of predicting vehicle counts by zip code, using various socio-economic and demographic features derived from IRS tax data. The analysis identified the Random Forest model as the most accurate and robust among the evaluated models, including XGBoost, Decision Tree, and HistGradientBoosting. In addition to the model performance, a detailed examination of the top features contributing to the model predictions provides valuable insights into the socio-economic factors influencing vehicle counts, particularly for electric vehicles (EVs).

### Key Feature Analysis

The top 10 features contributing to the model predictions include:

    Net Premium Tax Credit Amount: This feature represents the tax credit amount received by individuals under the Affordable Care Act. A higher tax credit may indicate lower income levels, suggesting a potential relationship between income and the ability to purchase vehicles, including EVs.

    Home Mortgage Interest Paid Amount: This feature reflects the interest paid on home mortgages, often correlating with homeownership and wealth. Areas with higher mortgage interest payments may have a higher concentration of middle to upper-income households, which can influence vehicle ownership patterns, including the adoption of EVs.

    Real Estate Taxes Amount: Real estate taxes are indicative of property values and, indirectly, wealth. Higher real estate taxes often correspond to higher property values, which may correlate with a higher prevalence of EV ownership due to greater financial capacity.

    Self-Employed Health Insurance Deduction Amount: This deduction is available to self-employed individuals who pay their health insurance premiums. It reflects the presence of self-employed individuals in a region, who may have different financial and vehicle ownership patterns compared to salaried employees.

    Salaries and Wages Amount: This feature directly indicates income levels from employment. Higher aggregate salaries and wages in a zip code suggest greater disposable income, which can facilitate the purchase of vehicles, including newer and potentially more expensive EVs.

    Limited State and Local Taxes: This feature represents the capped amount of state and local taxes that can be deducted from federal taxes. The deduction cap, especially in high-tax states, can impact disposable income and financial planning, potentially influencing vehicle purchasing decisions.

    Home Mortgage from Personal Seller Amount: This feature indicates transactions involving home mortgages from personal sellers. While less common, these transactions can indicate unique financial arrangements in certain areas, which might correlate with specific economic profiles affecting vehicle ownership.

    Individual Retirement Arrangement (IRA) Payments Amount: Contributions to IRAs reflect financial planning and the presence of disposable income. Areas with higher IRA payments may have wealthier populations, potentially increasing the likelihood of EV adoption due to greater financial resources.

    Self-Employed (Keogh) Retirement Plans Amount: Similar to IRAs, Keogh plans are retirement savings plans for self-employed individuals. Contributions to these plans suggest higher income and financial planning, which can correlate with higher rates of vehicle ownership, including EVs.

    Additional Medicare Tax Amount: This tax applies to high-income earners and indicates a population segment with substantial earnings. Higher additional Medicare tax payments correlate with higher income levels, potentially leading to increased EV adoption due to greater financial capacity.

### Implications and Socio-Economic Insights

The analysis of these top features reveals a consistent pattern: higher vehicle counts, particularly for EVs, are associated with higher income levels, wealth indicators (such as mortgage interest and real estate taxes), and financial planning (such as retirement contributions). These factors suggest that regions with greater financial capacity are more likely to adopt EVs, reflecting broader trends in consumer behavior and market accessibility.

The emphasis on features related to self-employment and additional taxes for high earners highlights the impact of income diversity on vehicle ownership patterns. The presence of deductions and credits for self-employed individuals suggests that entrepreneurial regions may exhibit unique vehicle ownership trends, possibly favoring flexibility and lower-cost vehicles or, conversely, reflecting higher disposable income that supports EV adoption.

## Conclusion
The Random Forest model's ability to incorporate and weigh these diverse socio-economic features underscores its robustness and suitability for this type of predictive analysis. The insights gained from the top features not only enhance our understanding of the factors driving vehicle ownership but also offer valuable guidance for policymakers and urban planners aiming to promote EV adoption. Future research could expand on these findings by incorporating additional data sources, such as environmental factors or public transportation availability, to provide a more comprehensive view of the determinants of vehicle ownership and EV adoption.

{{< pagebreak >}}
## Citations
Jain, A. (n.d.). Electric vehicle population [Data set]. Kaggle. Retrieved June 6, 2024, from https://www.kaggle.com/datasets/jainaru/electric-vehicle-population <br/><br/>

Internal Revenue Service. (n.d.). SOI tax stats - Individual income tax statistics - ZIP code data (SOI). Retrieved June 16, 2024, from https://www.irs.gov/statistics/soi-tax-stats-individual-income-tax-statistics-zip-code-data-soi

### 2024-08-06 code/processing-code/pre_process_tax_data.py

@@ -1,124 +1,0 @@
-from pandas import read_csv, set_option, merge
-from tqdm import tqdm # unnecessary dependency, adding as a meme 
-from numpy import mean
-from pyspark.sql import SparkSession
-set_option('display.max_columns', None)
-
-def data_dict_pre_process(filename: str):
-    strip_variable_whitespace_list = []
-    data_dict_df = read_csv(f'data/raw-data/{filename}.csv')
-    for idx, row in data_dict_df.iterrows():
-        variable = row['VARIABLE NAME'].strip()
-        strip_variable_whitespace_list.append(variable)
-    data_dict_df['VARIABLE NAME'] = strip_variable_whitespace_list
-
-    data_dict_df_variable_names = data_dict_df['VARIABLE NAME'].tolist()
-    
-    tax_qualitative_variables = data_dict_df_variable_names[:19]
-    format_tax_qualitative_variables = []
-    for code in tqdm(tax_qualitative_variables):
-        format_tax_qualitative_variables.append(code)
-    
-    tax_numerical_variables = data_dict_df_variable_names[19:]
-    total_count_return_variables = []
-    total_amount_return_variables = []
-    for code in tqdm(tax_numerical_variables):
-        if code.startswith('N'):
-            total_count_return_variables.append(code)
-        else:
-            total_amount_return_variables.append(code)
-
-    amount_return_variables_new = format_tax_qualitative_variables + total_amount_return_variables
-    count_return_variables_new = tax_qualitative_variables + total_count_return_variables
-    description_amount_data_dict_list = data_dict_df['DESCRIPTION'].loc[data_dict_df['VARIABLE NAME'].isin(amount_return_variables_new)].tolist()
-    description_count_data_dict_list = data_dict_df['DESCRIPTION'].loc[data_dict_df['VARIABLE NAME'].isin(count_return_variables_new)].tolist()
-
-    amount_data_dict_df = data_dict_df.loc[data_dict_df['VARIABLE NAME'].isin(amount_return_variables_new)]
-    count_data_dict_df = data_dict_df.loc[data_dict_df['VARIABLE NAME'].isin(count_return_variables_new)]
-    amount_data_dict_df.to_csv('data/intermediate-data/data_dict_variable_total_amount.csv', index=False)
[…]

### 2024-08-06 products/manuscript/Manuscript.qmd
update supplementary information

--- 2024-07-13 products/manuscript/Manuscript.qmd
@@
 ---
-title: "Example Manuscript Template for a Data Analysis Project"
-subtitle: ""
+title: "Using IRS Data to Predict Electric Vehicle Count"
 author: Sri Lakshmi Chundru, Collin Real, Joaquin Ramirez, Seth Harris
-date: "`r Sys.Date()`"
 format:
@@
 
-The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. 
-You don't need to have exactly these sections, but the content covering those sections should be addressed.
+## Abstract
+This study investigates the use of advanced machine learning models to predict vehicle count by zip code, utilizing a dataset enriched with socio-economic, demographic, and regional features.
+The primary aim is to accurately estimate vehicle counts, aiding in efficient resource distribution and urban planning.
+We implemented and evaluated several models, including Decision Tree, Random Forest, XGBoost, and HistGradientBoosting, focusing on their accuracy, interpretability, and computational efficiency.
 
-This uses MS Word as output format.
-[See here](https://quarto.org/docs/output-formats/ms-word.html) for more information.
-You can switch to other formats, like html or pdf.
-See [the Quarto documentation](https://quarto.org/) for other formats.
+Each model was trained and validated using cross-validation and a hold-out test set, with performance metrics such as Mean Absolute Error (MAE), Mean Squared Error (MSE), and R-squared used to assess their effectiveness.
+Among the models, the Random Forest model demonstrated superior performance, offering the best balance of predictive accuracy and model robustness.
+It effectively captured the complex interactions between the features and the target variable, outperforming the other models in terms of both accuracy and generalization.
 
-```{r, echo=FALSE, message=FALSE}
-# load a few R packages
-library(here)
-library(knitr)
-```
+The results highlight the Random Forest model's effectiveness in handling diverse and complex datasets, making it a valuable tool for predicting vehicle counts in different zip codes.
+This study provides critical insights into the practical applications of machine learning in urban planning and resource management, particularly in the context of transportation and infrastructure development.
 
--   https://www.google.com/publicdata/directory
--   https://www.data.gov/
--   https://github.com/awesomedata/awesome-public-datasets
--   https://registry.opendata.aws/
--   https://www.kaggle.com/datasets
+## Introduction
+The increasing adoption of electric vehicles (EVs) is a pivotal shift in the transportation sector, driven by growing environmental concerns, advancements in technology, and supportive government policies.
+Understanding the factors that influence the distribution of EVs is essential for developing targeted policies and infrastructure, such as charging stations and incentives.
+However, EV adoption is not uniform across regions; it varies significantly based on socio-economic and demographic factors.
+This study aims to explore these variations by predicting vehicle counts, particularly EVs, at the zip code level using a comprehensive dataset that combines electric vehicle registration data with individual income tax statistics from the Internal Revenue Service (IRS).
 
-- https://www.kaggle.com/datasets/jainaru/electric-vehicle-population
+The IRS data offers detailed insights into various financial attributes of taxpayers by zip code, including income levels, deductions, and credits.
+By integrating this data with EV registration data, we can analyze the relationship between financial capacity, tax-related benefits, and vehicle ownership patterns.
+Such an analysis is crucial for identifying the socio-economic determinants of EV adoption, which can inform policymakers and urban planners.
 
-# Summary/Abstract
-_Write a summary of your project._
+The main objectives of this study are twofold: first, to develop predictive models that accurately estimate vehicle counts by zip code, and second, to identify and quantify the impact of key socio-economic factors on these counts.
+We employed a variety of machine learning models, including Decision Tree, Random Forest, XGBoost, and HistGradientBoosting, to achieve these objectives.
+These models were chosen for their ability to handle complex, structured data and provide interpretable results.
 
-The dataset obtained from Kaggle provides detailed information about Battery Electric Vehicles (BEVs) and Plug-in Hybrid Electric Vehicles (PHEVs) registered through the Washington State Department of Licensing (DOL).
-It comprises 181,458 observations and 17 variables, encompassing key details such as Vehicle Identification Number (VIN), county, city, model year, electric range, and more.
+The findings of this study are expected to contribute valuable insights into the socio-economic factors driving EV adoption.
+By identifying the key predictors of vehicle counts, we can better understand the barriers to and facilitators of EV adoption in different communities.
+This knowledge is crucial for designing effective policies and interventions that promote sustainable transportation solutions.
+Furthermore, the study's methodology, which includes rigorous data cleaning, feature transformation, and model validation, provides a robust framework for future research in this area.
+
+## Methods
+In this study, we employed a variety of machine learning models to predict vehicle counts by zip code using socio-economic and demographic features extracted from IRS tax data.
+The data preparation, feature transformation, and model training processes are detailed below.
+
+#### Data Preparation
+The dataset was compiled by merging electric vehicle registration data with IRS tax data, focusing specifically on zip codes within the state of Washington.
+This geographical focus provided a diverse socio-economic landscape, enhancing the robustness of our analysis.
+The dataset included features such as 'Net premium tax credit amount,' 'Home mortgage interest paid amount,' 'Real estate taxes amount,' and several others related to income, tax credits, and deductions.
+
+To prepare the dataset for analysis, we first addressed data quality issues.
+We standardized column names using regular expressions, ensuring consistency across the merged dataset.
+This step was essential for preventing errors during data processing and analysis.
+We also applied data filtering to include only the relevant zip codes from Washington state, further refining our dataset for localized analysis.
+
+#### Feature Transformation
+Given that many of the predictor variables exhibited skewed distributions, we applied a log transformation to normalize these features.
+This transformation was particularly necessary for financial features, such as 'Salaries and wages amount' and 'Home mortgage interest paid amount,' which spanned several orders of magnitude.
+The transformation formula used was log⁡(x+1)log(x+1), which effectively handled zero values and stabilized variance across the dataset.
+Normalizing the data in this manner helped improve the performance and interpretability of our machine learning models by reducing the impact of outliers and facilitating a more uniform distribution of feature values.
+
+#### Model Development
+We implemented four different machine learning models: Decision Tree, Random Forest, XGBoost, and HistGradientBoosting.
+Each model was selected for its unique strengths in handling structured data and complex relationships between features.
+
+#### Model Training and Validation
+Each model was trained on the processed dataset, with training and validation sets created using cross-validation techniques.
+Cross-validation helped ensure that the models generalize well to unseen data, preventing overfitting.
+We evaluated model performance using metrics such as Root Mean Squared Error (RMSE) and R-squared, which provided insights into the accuracy and explanatory power of each model.
+
+Overall, the methods employed in this study—from data preparation and feature transformation to model development and validation—were designed to rigorously test and compare the predictive capabilities of multiple machine learning models.
+These methods provided a comprehensive framework for analyzing the socio-economic factors influencing vehicle counts by zip code, with a particular emphasis on the adoption of electric vehicles.
+
+## Results
+This section presents the comparative analysis of various machine learning models used to predict vehicle count by zip code.
+The models evaluated include XGBoost, Decision Tree, HistGradientBoosting, and Random Forest.
+The primary evaluation metrics considered were Root Mean Squared Error (RMSE) and R-squared, which provide insights into the accuracy and explanatory power of the models.
+
+### Model Performance Comparison
+#### XGBoost
+  - RMSE: 71.27 <br/>
+  - R-squared: 0.91 <br/>
+  - XGBoost demonstrated strong predictive capabilities, achieving a relatively low RMSE of 71.27.
+The model's R-squared value of 0.91 indicates that it explains 91% of the variance in the vehicle count data, suggesting good model fit and predictive accuracy.
+<br/>
+![XGBoost](../../results/output/model_results/XGBoost_Regressor.png)
+
+#### Decision Tree
+  - RMSE: 100.77<br/>
+  - R-squared: 0.83<br/>
+  -  The Decision Tree model had the highest RMSE among the evaluated models at 100.77, reflecting less precise predictions.
+Its R-squared value of 0.83 indicates that it explains 83% of the variance, making it the least accurate model in this comparison.
+![Decision Tree](../../results/output/model_results/Decision_Tree_Regressor.png)
+
+#### HistGradientBoosting
+  -  RMSE: 83.42<br/>
+  -  R-squared: 0.88<br/>
+  -  The HistGradientBoosting model offered a moderate performance with an RMSE of 83.42 and an R-squared value of 0.88.
+While better than the Decision Tree model, it still lagged behind XGBoost and Random Forest in terms of prediction accuracy and variance explanation.
+![HistGradientBoosting](../../results/output/model_results/HistGradientBoosting_Regressor.png)
+
+#### Random Forest
+  -  RMSE: 62.16<br/>
+  -  R-squared: 0.935<br/>
+  -  The Random Forest model outperformed all other models, achieving the lowest RMSE of 62.16 and the highest R-squared value of 0.935.
+These metrics indicate that the Random Forest model not only provided the most accurate predictions but also explained 93.5% of the variance in vehicle counts, making it the most reliable model in this analysis.
+![Random Forest](../../results/output/model_results/Random_Forest_Regressor.png)
+
+### Best Model Selection
+
+Based on the comparative analysis of the models, the Random Forest model emerged as the best performing model.
+It achieved the lowest RMSE and the highest R-squared, demonstrating superior accuracy and generalization capability compared to the other models.
+The Random Forest model's ensemble nature allows it to effectively capture complex patterns in the data, reduce overfitting through bootstrapping, and handle interactions between predictor variables more adeptly.
+These attributes collectively contributed to its superior performance, making it the most suitable choice for predicting vehicle counts by zip code in this study.
+
+The excellent performance of the Random Forest model suggests it is well-suited for applications requiring high prediction accuracy and robustness, particularly in urban planning and resource allocation scenarios where accurate vehicle count predictions are crucial.
+
+## Discussion
+The objective of this study was to develop machine learning models capable of predicting vehicle counts by zip code, using various socio-economic and demographic features derived from IRS tax data.
+The analysis identified the Random Forest model as the most accurate and robust among the evaluated models, including XGBoost, Decision Tree, and HistGradientBoosting.
+In addition to the model performance, a detailed examination of the top features contributing to the model predictions provides valuable insights into the socio-economic factors influencing vehicle counts, particularly for electric vehicles (EVs).
+
+### Key Feature Analysis
+
+The top 10 features contributing to the model predictions include:
+
+    Net Premium Tax Credit Amount: This feature represents the tax credit amount received by individuals under the Affordable Care Act.
+A higher tax credit may indicate lower income levels, suggesting a potential relationship between income and the ability to purchase vehicles, including EVs.
+
+    Home Mortgage Interest Paid Amount: This feature reflects the interest paid on home mortgages, often correlating with homeownership and wealth.
+Areas with higher mortgage interest payments may have a higher concentration of middle to upper-income households, which can influence vehicle ownership patterns, including the adoption of EVs.
+
+    Real Estate Taxes Amount: Real estate taxes are indicative of property values and, indirectly, wealth.
+Higher real estate taxes often correspond to higher property values, which may correlate with a higher prevalence of EV ownership due to greater financial capacity.
+
+    Self-Employed Health Insurance Deduction Amount: This deduction is available to self-employed individuals who pay their health insurance premiums.
+It reflects the presence of self-employed individuals in a region, who may have different financial and vehicle ownership patterns compared to salaried employees.
+
+    Salaries and Wages Amount: This feature directly indicates income levels from employment.
+Higher aggregate salaries and wages in a zip code suggest greater disposable income, which can facilitate the purchase of vehicles, including newer and potentially more expensive EVs.
+
+    Limited State and Local Taxes: This feature represents the capped amount of state and local taxes that can be deducted from federal taxes.
+The deduction cap, especially in high-tax states, can impact disposable income and financial planning, potentially influencing vehicle purchasing decisions.
+
+    Home Mortgage from Personal Seller Amount: This feature indicates transactions involving home mortgages from personal sellers.
+While less common, these transactions can indicate unique financial arrangements in certain areas, which might correlate with specific economic profiles affecting vehicle ownership.
+
+    Individual Retirement Arrangement (IRA) Payments Amount: Contributions to IRAs reflect financial planning and the presence of disposable income.
+Areas with higher IRA payments may have wealthier populations, potentially increasing the likelihood of EV adoption due to greater financial resources.
+
+    Self-Employed (Keogh) Retirement Plans Amount: Similar to IRAs, Keogh plans are retirement savings plans for self-employed individuals.
+Contributions to these plans suggest higher income and financial planning, which can correlate with higher rates of vehicle ownership, including EVs.
+
+    Additional Medicare Tax Amount: This tax applies to high-income earners and indicates a population segment with substantial earnings.
+Higher additional Medicare tax payments correlate with higher income levels, potentially leading to increased EV adoption due to greater financial capacity.
+
+### Implications and Socio-Economic Insights
+
+The analysis of these top features reveals a consistent pattern: higher vehicle counts, particularly for EVs, are associated with higher income levels, wealth indicators (such as mortgage interest and real estate taxes), and financial planning (such as retirement contributions).
+These factors suggest that regions with greater financial capacity are more likely to adopt EVs, reflecting broader trends in consumer behavior and market accessibility.
+
+The emphasis on features related to self-employment and additional taxes for high earners highlights the impact of income diversity on vehicle ownership patterns.
+The presence of deductions and credits for self-employed individuals suggests that entrepreneurial regions may exhibit unique vehicle ownership trends, possibly favoring flexibility and lower-cost vehicles or, conversely, reflecting higher disposable income that supports EV adoption.
+
+## Conclusion
+The Random Forest model's ability to incorporate and weigh these diverse socio-economic features underscores its robustness and suitability for this type of predictive analysis.
+The insights gained from the top features not only enhance our understanding of the factors driving vehicle ownership but also offer valuable guidance for policymakers and urban planners aiming to promote EV adoption.
+Future research could expand on these findings by incorporating additional data sources, such as environmental factors or public transportation availability, to provide a more comprehensive view of the determinants of vehicle ownership and EV adoption.
 
 {{< pagebreak >}}
+## Citations
+Jain, A. (n.d.). Electric vehicle population [Data set]. Kaggle. Retrieved June 6, 2024, from https://www.kaggle.com/datasets/jainaru/electric-vehicle-population <br/><br/>
 
-# Introduction
-
-**- (a) Briefly describe what the data is, how it was collected, where you will get (or got) it from.
-How many observations do you have, what was measured?
-Anything else important to report about the data?**
-
-The dataset obtained from Kaggle provides detailed information about Battery Electric Vehicles (BEVs) and Plug-in Hybrid Electric Vehicles (PHEVs) registered through the Washington State Department of Licensing (DOL).
-It comprises 181,458 observations and 17 variables, encompassing key details such as Vehicle Identification Number (VIN), county, city, model year, electric range, and more.
-
-**- (b) At this stage you are not required to already have and show the data, but if you do, even better.
-Then add a few lines of code which load the data and using some of the commands you learned about, provide summary descriptions of the data.**
-```{r}
-data_location <- here::here("data","raw-data","Electric_Vehicle_Population_Data.csv")
-data <- read.csv(data_location)
-str(data)
-head(data)
-summary(data)
-colSums(is.na(data))
-```
-
-**- (c) Explain the question you want to answer using the data.
-What will be your outcome(s) of interest (if any)?
-What (if any) specific predictors will you focus on?
-What relations/patterns are you looking for in the data?**
-
-Question: What insights can be gleaned from the dataset regarding the electric vehicle landscape ?
-
-## Objectives:
-
-- Understand Geographic Distribution: Determine the distribution of electric vehicles across different counties and cities in Washington State.
-- Identify Popular Models: Identify the most commonly registered electric vehicle makes and models.
-- Analyze Electric Range: Investigate the distribution of electric range among registered electric vehicles.-
-- Explore Clean Alternative Fuel Vehicle Eligibility: Examine the eligibility for clean alternative fuel vehicle incentives among different vehicle types.
-
-## Outcomes of Interest
-- Geographic Distribution: The distribution of electric vehicles across counties and cities in Washington State.
-- Popular Models: The most frequently registered electric vehicle makes and models.
-- Electric Range Analysis: The range distribution and average electric range of registered electric vehicles.
-- Clean Alternative Fuel Vehicle Eligibility: The proportion of vehicles eligible for clean alternative fuel vehicle incentives.
-
-## Predictors
-- Geographic Variables: County, City, Postal Code.
-- Vehicle Attributes: Model Year, Make, Model, Electric Vehicle Type, Electric Range.
-
-## Relations/Patterns
-- Geographic Distribution: Analyzing clusters of electric vehicles in specific regions.
-- Popular Models: Identifying dominant electric vehicle makes and models.
-- Electric Range Analysis: Understanding the distribution of electric ranges and potential implications for consumer preferences.
-- Clean Alternative Fuel Vehicle Eligibility: Examining eligibility patterns and potential factors influencing eligibility.
-- Correlations: Investigating any associations between electric vehicle characteristics and legislative districts or census tracts, which may offer insights into policy implications and consumer behavior.
-
-**(d) As much as you know, suggest how you will analyze it. At this stage in the course, we haven’t covered analysis approaches yet, so you can keep things vague and non-technical here.**
-
-- Descriptive Analysis: Summarize the dataset to understand its contents, including summary statistics and frequency tables.
-
--Visualization: Use plots and charts to visualize aspects like geographic distribution, electric range, and vehicle types.
-
-- Comparative Analysis: Compare different groups within the dataset, such as BEVs vs PHEVs or eligibility for incentives by vehicle type.
-
-- Correlation Analysis: Explore potential relationships between variables, such as electric range and base MSRP.
-
-- Geospatial Analysis (optional): Use R packages to analyze spatial patterns in the data.
-
-- Exploratory Data Analysis (EDA): Engage in exploratory analysis to identify interesting patterns or trends.
-
-**(e) You are allowed, but not yet required, to provide background information for the question you plan to answer.
-For instance you can describe why it’s an interesting question, who else has done similar analyses, how your analysis will be new/different, etc.
-Similar to what you read in an introduction to a research paper.
-For the final report, you’ll need these parts.
-For part 1, they are not required, but you are welcome to already write down some of that.
-**
-
-## Primary Research Question: What insights can be gleaned from the dataset regarding the electric vehicle landscape in Washington State?
-
-Sub-questions:
-
--  How are electric vehicles distributed across different counties, cities, and States?
--  Which EV makes and models are most popular?
--  What is the range of electric vehicles in terms of electric range and base MSRP?
--  How does eligibility for clean alternative fuel vehicle incentives vary among different vehicle types?
--  Are there any significant correlations between EV characteristics and geographic or legislative factors?
-
-**(d) Eventually, for your final report, what you write for this part will go into different sections of the full report.
-Some will go into the introduction, some in the methods section.
-You can already place these items there, or for now just write them as a single section.**
-
-**Introduction**
-Electric vehicles (EVs), including Battery Electric Vehicles (BEVs) and Plug-in Hybrid Electric Vehicles (PHEVs), are crucial in the global effort to reduce greenhouse gas emissions and dependence on fossil fuels.
-This dataset, sourced from the Washington State Department of Licensing (DOL) and obtained from Kaggle, provides a comprehensive overview of registered electric vehicles in Washington State.
-Analyzing this data offers valuable insights into the adoption patterns and characteristics of EVs, informing policies and infrastructure development to further promote sustainable transportation.
-
-**Objectives:**
-
-**Geographic Distribution:** Determine the distribution of electric vehicles across different counties and cities.
-**Popular Models:** Identify the most commonly registered EV makes and models.
-**Electric Range Analysis:** Investigate the distribution of electric range among registered EVs.
-**CAFV Eligibility:** Examine the eligibility for clean alternative fuel vehicle incentives among different vehicle types.
-**Correlation Analysis:** Explore potential correlations between EV characteristics and geographic or legislative factors.
-
-**Methods**
-
-**Data Cleaning and Preparation:**
-
-- Address missing values and ensure correct data types.
-- Transform variables as necessary for analysis.
-
-To ensure the dataset is ready for analysis, the following steps were undertaken:
-
-**1. Data Import and Initial Inspection:**
-
-- The dataset Electric_Vehicle_Population_Data.csv was imported.
-- Summary statistics and data types were displayed to understand the initial structure and content of the data.
-- UPDATE (6.17.2024) - Downloaded IRS tax data to join with the EV data by zip code
-  - Tax data provides many financial variables that we will include in our models (i.e. can we predict the amount of EVs per zip code based off tax data?)
-- Added a geojson file to draw zip codes on a map, will use for visualizations but still figuring out how to elegantly filter the json
-
-**2. Checking for Duplicates:**
-
-- Identified and counted the number of duplicate rows in the dataset to avoid redundant information.
-- Result: num_duplicates duplicates were found.
-
-**3. Counting Rows with Zero MSRP:**
-
-- Counted the number of rows where the Base.MSRP (Manufacturer's Suggested Retail Price) is zero to identify incomplete pricing data.
-- Result: count_zero_msrp rows with zero MSRP were identified.
-
-**4. Checking for Null Values:**
-
-- Calculated the total number of null (NA) values in the dataset to address missing data.
-- Result: total_na null values were found.
-
-**5. Removing Null Values and Identifying Outliers:**
-
-- Filtered out rows with null values to ensure data completeness.
-- Identified the maximum Base.MSRP for each electric vehicle type to spot potential outliers.
-- Result:The maximum MSRP for each vehicle type was summarized.
-
-**6. Filtering Out Outliers and Zero MSRP:**
-
-Removed the row with an extreme outlier Base.MSRP of 845,000 to prevent distortion in the analysis.
-Also filtered out rows where Base.MSRP is zero to ensure only valid pricing data is considered.
-By performing these steps, the dataset was cleaned and prepared, ensuring it is more accurate and reliable for subsequent analysis.
-
-**Descriptive Statistics:**
-
-- Compute summary statistics (mean, median, mode, standard deviation) for numerical variables.
-- Generate frequency tables for categorical variables.
-
-**Boxplot - Price Distribution vs. Electric Vehicle Type:**
-
-This boxplot compares the price distribution of Battery Electric Vehicles (BEV) and Plug-in Hybrid Electric Vehicles (PHEV).
-It shows that BEVs have a wider price range with a higher median price compared to PHEVs, which have a more compact price distribution with some outliers.
-
-**Boxplot - Price Distribution vs. Electric Vehicle Type (Clean Alternative Fuel Vehicle Eligible):**
-
-This boxplot illustrates the price distribution based on eligibility for Clean Alternative Fuel Vehicle incentives.
-Vehicles eligible for the incentive tend to have a higher price range compared to those that are not eligible due to low battery range.
-
-**Boxplot - Price Distribution vs. Electric Vehicle Type (Drive Range):**
-
-This boxplot shows the distribution of drive range for clean alternative fuel vehicle eligible and non-eligible vehicles.
-Eligible vehicles have a wider and higher range compared to non-eligible vehicles, which have a limited drive range.
-
-**Scatterplot - Electric Vehicle Range Growth over Time:**
-
-This scatterplot depicts the growth in electric vehicle range over time.
-There is a clear trend showing that the electric range of vehicles has generally increased over the years.
-
-**Boxplot - Price Distribution by Brand:**
-
-This boxplot compares the price distribution of electric vehicles by different brands.
-Tesla has the highest price range, followed by Porsche. Brands like KIA and Chrysler have lower price distributions.
-
-**Scatterplot - Price vs. Range Correlation:**
-
-This scatterplot shows the correlation between the vehicle price and its electric range.
-There is a positive correlation where higher-priced vehicles tend to have a longer electric range.
-These plots provide insights into the price and range variations of electric vehicles based on type, eligibility for incentives, brand, and over time.
-
-**Data Visualization:**
-
-- Create bar charts, histograms, and pie charts to visualize distributions.
-- Use scatter plots and box plots to explore relationships between variables.
-- Generate geographic maps to visualize EV distribution across the state.
-
-**Comparative Analysis:**
-
-- Compare BEVs and PHEVs in terms of electric range and base MSRP.
-- Analyze differences in CAFV eligibility across makes and models.
-
-**Correlation Analysis:**
-
-- Calculate correlation coefficients for numerical variables.
-- Use cross-tabulations to explore relationships between categorical and numerical variables.
-
-**Geospatial Analysis (if applicable):**
-
-- Use geographic mapping tools to identify clusters or patterns in vehicle registrations.
-
-**Exploratory Data Analysis (EDA):**
-
--Perform in-depth analysis to identify trends, patterns, or outliers.
--Use advanced visualization techniques for deeper insights.
-
-**Statistical Model**
-- First iteration of models
-- Produced a few statistically significant tax metrics for predicting the number of EVs in a given county
-- Washington demographics datasets could further enhance our analysis
-
-**Conclusion:**
-
-By analyzing the electric vehicle registration data for Washington State, this project aims to uncover significant patterns and trends that can inform policy decisions and infrastructure planning.
-The insights gained from this analysis will contribute to the broader understanding of EV adoption and support the transition towards a more sustainable transportation future.
-
-{{< pagebreak >}}
+Internal Revenue Service.
+(n.d.).
+SOI tax stats - Individual income tax statistics - ZIP code data (SOI).
+Retrieved June 16, 2024, from https://www.irs.gov/statistics/soi-tax-stats-individual-income-tax-statistics-zip-code-data-soi

### 2024-08-06 products/manuscript/supplement/Supplementary-Material.qmd
update supplementary information

@@
 subtitle: ""
-author: Andreas Handel
-date: "`r Sys.Date()`"
+author: Sri Lakshmi Chundru, Collin Real, Joaquin Ramirez, Seth Harris
 format:
-  pdf:
+  docx:
     toc: false
@@
     highlight-style: github
-#bibliography: ../../../assets/dataanalysis-references.bib
-csl: ../../../assets/vancouver.csl
+# bibliography: ../../assets/dataanalysis-references.bib
+# csl: ../../assets/apa.csl
 ---
-
-```{r, echo=FALSE, message=FALSE}
-# load a few R packages
-library(here)
-library(knitr)
-```
-
-This shows some materials that could go into a supplementary file.
-Often you want/need references here too.
-You can use the same reference bib file for this and the main text (as done here) or have separate bib files.
-
-For illustrative purposes, I'm doing the supplement as pdf.
-For this to work, you need a (La)TeX system installed.
-It's easy.
-Just follow [these steps](https://quarto.org/docs/output-formats/pdf-basics.html).
-
-Of course you would choose the format based on needs.
-
-I'm also using a different style for the references here.
-(vancouver vs apa in the main manuscript).
-Usually one would have the formatting of the references the same in those two documents, but I want to illustrate how easy it is to switch reference formatting styles, you just need to get the right CSL file and specify it in the YAML header.
-We could also have a seperate reference bibtext (`.bib`) file, but here we are using the same.
 
@@
 
-A quick overview of what readers can find in the supplement.
+An overview of the correct process to generate the final outputs as well as some exploratory analysis.
 
-# Code and file information
+# Code Information
 
-Explain here what each code/file is and does, and in which order (if any) users need to run thing to reproduce everything.
-Essentially, give a full set of instructions to re-generate everything.
+**Run analysis in the following order.**
 
-{{< pagebreak >}}
+1. Pre-Processing File (Quarto): <br/>
+/code/processing-code/pre_process_tax_data.qmd - preprocesses the tax/ev data and merges them into one dataset.
+
+2. Exploratory Data Analysis (Quarto): <br/>
+/code/eda-code/eda-ev-dataset.qmd - exploratory analysis of the electric vehicle dataset <br/>
+/code/eda-code/eda-ev-tax.qmd - exploratory analysis of the tax dataset <br/>
+
+3. Predictive models (Quarto): <br/>
+/code/analysis-code/statistical_analysis.qmd - builds multiple predictive models, visualizes, and compares them.
 
@@
 
-Often, the main manuscript only allows for an overview description of the methods.
-Use the supplement to describe all your methods, models and approaches in a lot of detail.
-Reference specific parts of your code as needed.
+#### Decision Tree Regressor:
+A basic model that splits the data into subsets based on feature values, creating a tree structure.
+We set hyperparameters such as the maximum depth and minimum samples per leaf to control the model's complexity and prevent overfitting.
 
-{{< pagebreak >}}
+#### Random Forest Regressor:
+An ensemble learning method that constructs multiple decision trees and aggregates their outputs to improve prediction accuracy.
+We configured key parameters like the number of trees (n_estimators) and the maximum depth, optimizing for a balance between bias and variance.
+
+#### XGBoost Regressor:
+A gradient boosting model known for its efficiency and performance.
+XGBoost builds trees sequentially, with each new tree aiming to correct errors made by the previous ones.
+We fine-tuned parameters such as the learning rate (eta), maximum depth, and regularization terms (lambda and alpha) to enhance model performance.
+
+#### HistGradientBoosting Regressor:
+A histogram-based implementation of gradient boosting that is particularly efficient for large datasets.
+This model discretizes continuous features into bins, reducing memory usage and speeding up training.
+We set parameters including the number of iterations (max_iter), learning rate, and number of bins (max_bins) to optimize the model.
 
 # Additional results
+## Log-Transformed Histogram Examples
+![SE Health Insurance Deduction](../../../results/figures/histograms/Self-employed%20health%20insurance%20deduction%20amount_histogram.png)
 
-Show additional results here. Those can be some useful exploratory/descriptive figures or tables, or results from additional analyses that didn't make it into the main text.
+![Salary & Wages Amount](../../../results/figures/histograms/Salaries%20and%20wages%20amount_histogram.png)
 
-## Example additional result
+![Net Premium Tax Credit Amount](../../../results/figures/histograms/Net%20premium%20tax%20credit%20amount_histogram.png)
 
-@tbl-resulttable1 shows an additional table summarizing a model fit.
+## Log-Transformed Boxplot Examples
+![SE Health Insurance Deduction](../../../results/figures/boxplots/Self-employed%20health%20insurance%20deduction%20amount_boxplot.png)
 
-```{r}
-#| label: tbl-resulttable1
-#| tbl-cap: "Another fit table."
-#| echo: FALSE
-resulttable1 = readRDS(here("results","tables","resulttable1.rds"))
-knitr::kable(resulttable1)
-```
+![Salary & Wages Amount](../../../results/figures/boxplots/Salaries%20and%20wages%20amount_boxplot.png)
 
-@fig-result2 shows a scatterplot figure produced by one of the R scripts.
+![Net Premium Tax Credit Amount](../../../results/figures/boxplots/Net%20premium%20tax%20credit%20amount_boxplot.png)
 
-```{r}
-#| label: fig-result2
-#| fig-cap: "Height and weight."
-#| echo: FALSE
-knitr::include_graphics(here("results","figures","height-weight.png"))
-```
+## Model Comparison - MAE
+!['MAE'](../../../results/output/model_comparison/MAE.png)
 
-{{< pagebreak >}}
+## Model Comparison - MSE
+!['MSE'](../../../results/output/model_comparison/MSE.png)
 
-# Discussion
+## Model Comparison - R-squared
+!['R-squared'](../../../results/output/model_comparison/R-squared.png)
 
-Any additional discussion regarding the supplementary material/findings.
-
-These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.
-
-{{< pagebreak >}}
+## Model Comparison - RMSE
+!['RMSE'](../../../results/output/model_comparison/RMSE.png)
 
 # References
+- Jain, A. (n.d.). Electric vehicle population [Data set]. Kaggle. Retrieved June 6, 2024, from https://www.kaggle.com/datasets/jainaru/electric-vehicle-population <br/>
+
+- Internal Revenue Service.
+(n.d.).
+SOI tax stats - Individual income tax statistics - ZIP code data (SOI).
+Retrieved June 16, 2024, from https://www.irs.gov/statistics/soi-tax-stats-individual-income-tax-statistics-zip-code-data-soi

### 2024-08-06 products/manuscript/Manuscript.qmd
add citations to manuscript

@@
 However, EV adoption is not uniform across regions; it varies significantly based on socio-economic and demographic factors.
-This study aims to explore these variations by predicting vehicle counts, particularly EVs, at the zip code level using a comprehensive dataset that combines electric vehicle registration data with individual income tax statistics from the Internal Revenue Service (IRS).
+This study aims to explore these variations by predicting vehicle counts, particularly EVs, at the zip code level using a comprehensive dataset that combines electric vehicle registration data @kaggle with individual income tax statistics from the Internal Revenue Service (IRS) @irs.
 
@@
 ## Citations
-Jain, A. (n.d.). Electric vehicle population [Data set]. Kaggle. Retrieved June 6, 2024, from https://www.kaggle.com/datasets/jainaru/electric-vehicle-population <br/><br/>
+Jain, A. (n.d.). Electric vehicle population [Data set]. Kaggle. Retrieved June 6, 2024, from https://www.kaggle.com/datasets/jainaru/electric-vehicle-population <br/>
 

### 2024-08-06 code/analysis-code/statistical_analysis.qmd
add scatterplots and update Manuscript

@@
 from sklearn.model_selection import train_test_split
+from sklearn.preprocessing import StandardScaler
 import re
@@
 y_grouped = merged_data['Vehicle Count']
+
+scalar = StandardScaler()
+scale = scalar.fit_transform(X_grouped)
+print(scale)
 
@@
 
-rand_for_feature_importance_list = rand_for_feature_importance['Feature'].head(10).tolist()
+rand_for_feature_importance_list = xgboost_feature_importance['Feature'].head(10).tolist()
 print(rand_for_feature_importance_list)

### 2024-08-06 code/eda-code/eda-ev-tax.qmd
add scatterplots and update Manuscript

@@
 X_train_grouped, X_test_grouped, y_train_grouped, y_test_grouped = train_test_split(
-    X_grouped, y_grouped, test_size=0.3, random_state=42)
+    X_grouped, y_grouped, test_size=0.2, random_state=42)
+
+merged_data.to_csv('test.csv', index=False)
+```
+
+```{python}
+import matplotlib.pyplot as plt
+
+# Scatter plots to understand relationship between predictors and independent variable
+for i, feature in enumerate(features_to_use):
+    plt.figure(figsize=(10,5))
+    plt.scatter(X_grouped[feature], y_grouped, c="lightblue", edgecolors="blue")
+    plt.title(f'{feature} vs Vehicle Count')
+    plt.xlabel(feature)
+    plt.ylabel('Vehicle Count')
+    # plt.grid(True)
+    plt.savefig(f'../../results/figures/scatterplots/{feature}_scatterplot.png')
 
@@
 # Preprocess and log transform the training and test datasets
-X_train_transformed = preprocess_and_log_transform_data(X_train_grouped)
-X_test_transformed = preprocess_and_log_transform_data(X_test_grouped)
+X_transformed = preprocess_and_log_transform_data(X_grouped)
 
@@
     for feature in data.columns:
-        fig, ax = plt.subplots(1, 2, figsize=(10,10))
+        fig, ax = plt.subplots(1, 2, figsize=(10,5))
         ax1 = plt.subplot(2,1,1)
@@
         ax2 = sns.histplot(data1[feature], kde=True, bins=50)
-        ax1.title.set_text(f'Histogram of {feature}')
-        ax2.title.set_text(f'Log-Transformed Histogram of {feature}')
+        ax1.title.set_text(f'Original vs. Log-Transformed: {feature}')
+        # ax2.title.set_text(f'Log-Transformed Histogram of {feature}')
         plt.xlabel(feature)
@@
 # Plot the histograms for the log-transformed features
-plot_transformed_feature_histograms(X_train_grouped, X_train_transformed)
+plot_transformed_feature_histograms(X_grouped, X_transformed)
 ```
@@
         ax2 = sns.boxplot(data1[feature])
-        ax1.title.set_text(f'Boxplot of {feature}')
-        ax2.title.set_text(f'Log-Transformed Boxplot of {feature}')
+        ax1.title.set_text(f'Original vs. Log-Transformed Boxplot of {feature}')
+        # ax2.title.set_text(f'Log-Transformed Boxplot of {feature}')
         plt.xlabel(feature)
@@
 # Plot the boxplots for the log-transformed features
-plot_transformed_feature_boxplots(X_train_grouped, X_train_transformed)
+plot_transformed_feature_boxplots(X_grouped, X_transformed)
 ```

### 2024-08-06 products/manuscript/Manuscript.qmd
add scatterplots and update Manuscript

@@
 
-The results highlight the Random Forest model's effectiveness in handling diverse and complex datasets, making it a valuable tool for predicting vehicle counts in different zip codes.
+The results highlight the XGBoost model's effectiveness in handling diverse and complex datasets, making it a valuable tool for predicting vehicle counts in different zip codes.
 This study provides critical insights into the practical applications of machine learning in urban planning and resource management, particularly in the context of transportation and infrastructure development.
@@
 
-#### Data Preparation
+### Data Preparation & Exploratory Analysis
 The dataset was compiled by merging electric vehicle registration data with IRS tax data, focusing specifically on zip codes within the state of Washington.
@@
 
-#### Feature Transformation
+#### Scatterplots - Predictor's vs. Vehicle Count Relationship
+![SE Health Insurance Deduction Scatterplot](../../results/figures/scatterplots/Self-employed%20health%20insurance%20deduction%20amount_scatterplot.png)
+
+### Feature Transformation
 Given that many of the predictor variables exhibited skewed distributions, we applied a log transformation to normalize these features.
@@
 
-#### Model Development
+#### Log-Transform Histogram
+![SE Health Insurance Deduction](../../results/figures/histograms/Self-employed%20health%20insurance%20deduction%20amount_histogram.png)
+
+#### Log-Transform Boxplot
+![SE Health Insurance Deduction Boxplot](../../results/figures/boxplots/Self-employed%20health%20insurance%20deduction%20amount_boxplot.png)
+
+### Model Development
 We implemented four different machine learning models: Decision Tree, Random Forest, XGBoost, and HistGradientBoosting.
@@
 
-#### Model Training and Validation
+### Model Training and Validation
 Each model was trained on the processed dataset, with training and validation sets created using cross-validation techniques.
@@
 #### XGBoost
-  - RMSE: 71.27 <br/>
-  - R-squared: 0.91 <br/>
+  - RMSE: 69.21 <br/>
+  - R-squared: 0.892 <br/>
   - XGBoost demonstrated strong predictive capabilities, achieving a relatively low RMSE of 71.27.
-The model's R-squared value of 0.91 indicates that it explains 91% of the variance in the vehicle count data, suggesting good model fit and predictive accuracy.
+The model's R-squared value of 0.91 indicates that it explains 89.2% of the variance in the vehicle count data, suggesting good model fit and predictive accuracy.
 <br/>
@@
 #### Decision Tree
-  - RMSE: 100.77<br/>
-  - R-squared: 0.83<br/>
-  -  The Decision Tree model had the highest RMSE among the evaluated models at 100.77, reflecting less precise predictions.
-Its R-squared value of 0.83 indicates that it explains 83% of the variance, making it the least accurate model in this comparison.
+  - RMSE: 79.61<br/>
+  - R-squared: 0.857<br/>
+  -  The Decision Tree model had the highest RMSE among the evaluated models at 79.61, reflecting less precise predictions.
+Its R-squared value of 0.857 indicates that it explains 85.7% of the variance, making it the least accurate model in this comparison.
 ![Decision Tree](../../results/output/model_results/Decision_Tree_Regressor.png)
@@
 #### HistGradientBoosting
-  -  RMSE: 83.42<br/>
-  -  R-squared: 0.88<br/>
-  -  The HistGradientBoosting model offered a moderate performance with an RMSE of 83.42 and an R-squared value of 0.88.
+  -  RMSE: 76.96<br/>
+  -  R-squared: 0.867<br/>
+  -  The HistGradientBoosting model offered a moderate performance with an RMSE of 76.96 and an R-squared value of 0.867.
 While better than the Decision Tree model, it still lagged behind XGBoost and Random Forest in terms of prediction accuracy and variance explanation.
@@
 #### Random Forest
-  -  RMSE: 62.16<br/>
-  -  R-squared: 0.935<br/>
-  -  The Random Forest model outperformed all other models, achieving the lowest RMSE of 62.16 and the highest R-squared value of 0.935.
-These metrics indicate that the Random Forest model not only provided the most accurate predictions but also explained 93.5% of the variance in vehicle counts, making it the most reliable model in this analysis.
+  -  RMSE: 71.30<br/>
+  -  R-squared: 0.885<br/>
+  -  The Random Forest model outperformed all other models, achieving the lowest RMSE of 71.30 and the highest R-squared value of 0.885.
+These metrics indicate that the Random Forest model not only provided the most accurate predictions but also explained 88.5% of the variance in vehicle counts, making it the most reliable model in this analysis.
 ![Random Forest](../../results/output/model_results/Random_Forest_Regressor.png)
@@
 
-Based on the comparative analysis of the models, the Random Forest model emerged as the best performing model.
-It achieved the lowest RMSE and the highest R-squared, demonstrating superior accuracy and generalization capability compared to the other models.
-The Random Forest model's ensemble nature allows it to effectively capture complex patterns in the data, reduce overfitting through bootstrapping, and handle interactions between predictor variables more adeptly.
-These attributes collectively contributed to its superior performance, making it the most suitable choice for predicting vehicle counts by zip code in this study.
-
-The excellent performance of the Random Forest model suggests it is well-suited for applications requiring high prediction accuracy and robustness, particularly in urban planning and resource allocation scenarios where accurate vehicle count predictions are crucial.
+The XGBoost model emerged as the best performer in predicting vehicle counts by zip code, primarily due to its advanced boosting techniques and the ability to handle various data complexities effectively.
+The combination of these features makes XGBoost a powerful and versatile tool for predictive modeling.
+Its ability to handle complex, structured data, prevent overfitting, and efficiently process large datasets, while offering detailed feature importance analysis, made it the top performer in this study.
+The model's superior performance in terms of low RMSE and high R-squared values underscores its effectiveness in accurately predicting vehicle counts by zip code, providing valuable insights for urban planning and policy-making.
 
 ## Discussion
-The objective of this study was to develop machine learning models capable of predicting vehicle counts by zip code, using various socio-economic and demographic features derived from IRS tax data.
-The analysis identified the Random Forest model as the most accurate and robust among the evaluated models, including XGBoost, Decision Tree, and HistGradientBoosting.
-In addition to the model performance, a detailed examination of the top features contributing to the model predictions provides valuable insights into the socio-economic factors influencing vehicle counts, particularly for electric vehicles (EVs).
+The top 10 features identified as the most significant predictors of vehicle counts by zip code offer insights into the socio-economic and financial characteristics of the population.
+These features provide a nuanced understanding of the factors that potentially influence vehicle ownership, particularly the adoption of electric vehicles (EVs).
+Here's a detailed discussion of each feature and its relevance:
 
-### Key Feature Analysis
+## Variable Importance - XGBoost
+!['XGBoost'](../../results/output/variable_importance/XGBoost.png)
 
-The top 10 features contributing to the model predictions include:
+1. Self-Employed Health Insurance Deduction Amount
+This feature represents the amount self-employed individuals can deduct for health insurance premiums.
+It is significant as it indicates the presence and financial activities of self-employed individuals within a zip code.
+Self-employed individuals often have variable incomes and unique financial situations, which may influence their ability to purchase and maintain vehicles.
+This deduction could correlate with areas where self-employment is prevalent, possibly reflecting regions with more entrepreneurial activity and, potentially, higher disposable incomes.
 
-    Net Premium Tax Credit Amount: This feature represents the tax credit amount received by individuals under the Affordable Care Act.
-A higher tax credit may indicate lower income levels, suggesting a potential relationship between income and the ability to purchase vehicles, including EVs.
+2. Individual Retirement Arrangement Payments Amount
+Contributions to Individual Retirement Arrangements (IRAs) reflect financial planning and savings behavior.
+Higher contributions suggest higher disposable income and financial stability, factors that are positively associated with the ability to purchase vehicles, including more expensive options like EVs.
+This feature indicates regions where residents prioritize long-term financial health, which may correlate with the capacity to invest in newer, environmentally friendly technologies.
 
-    Home Mortgage Interest Paid Amount: This feature reflects the interest paid on home mortgages, often correlating with homeownership and wealth.
-Areas with higher mortgage interest payments may have a higher concentration of middle to upper-income households, which can influence vehicle ownership patterns, including the adoption of EVs.
+3. Home Mortgage from Personal Seller Amount
+This feature involves transactions where the home mortgage is obtained directly from a personal seller rather than a traditional financial institution.
+It indicates non-standard financial arrangements, possibly reflecting unique housing markets or economic conditions in certain areas.
+The relevance of this feature could be tied to the accessibility and flexibility in financial transactions within these regions, potentially influencing overall economic activity, including vehicle purchases.
 
-    Real Estate Taxes Amount: Real estate taxes are indicative of property values and, indirectly, wealth.
-Higher real estate taxes often correspond to higher property values, which may correlate with a higher prevalence of EV ownership due to greater financial capacity.
+4. Salaries and Wages Amount
+A direct indicator of income, this feature represents the total salaries and wages reported within a zip code.
+It is a critical predictor as higher salaries and wages generally correlate with greater financial capacity to purchase and maintain vehicles.
+This feature strongly influences vehicle ownership levels, including the adoption of EVs, which may have higher upfront costs but offer long-term savings.
 
-    Self-Employed Health Insurance Deduction Amount: This deduction is available to self-employed individuals who pay their health insurance premiums.
-It reflects the presence of self-employed individuals in a region, who may have different financial and vehicle ownership patterns compared to salaried employees.
+5. Home Mortgage Interest Paid Amount
+The amount of interest paid on home mortgages is indicative of both property ownership and the financial burden of maintaining such properties.
+Higher mortgage interest payments suggest higher home values and, by extension, potentially higher wealth levels.
+This wealth is a significant factor in vehicle ownership, as individuals with more expensive homes often have higher disposable incomes, enabling them to invest in additional vehicles or more luxurious options, including EVs.
 
-    Salaries and Wages Amount: This feature directly indicates income levels from employment.
-Higher aggregate salaries and wages in a zip code suggest greater disposable income, which can facilitate the purchase of vehicles, including newer and potentially more expensive EVs.
+6. Self-Employment Tax Amount
+This feature captures the amount of self-employment tax paid, which directly correlates with self-employment income.
+Higher self-employment tax payments indicate successful businesses or high-earning self-employed individuals.
+The economic health of these individuals likely contributes to their ability to afford vehicles, making this a significant predictor of vehicle counts.
 
-    Limited State and Local Taxes: This feature represents the capped amount of state and local taxes that can be deducted from federal taxes.
-The deduction cap, especially in high-tax states, can impact disposable income and financial planning, potentially influencing vehicle purchasing decisions.
+7. Self-Employed (Keogh) Retirement Plans Amount
+Contributions to Keogh plans reflect savings set aside by self-employed individuals for retirement.
+Like IRA contributions, this feature suggests a focus on long-term financial security, indicative of higher income levels and financial planning.
+This financial stability is associated with greater capacity for discretionary spending, including vehicle purchases.
 
-    Home Mortgage from Personal Seller Amount: This feature indicates transactions involving home mortgages from personal sellers.
-While less common, these transactions can indicate unique financial arrangements in certain areas, which might correlate with specific economic profiles affecting vehicle ownership.
+8. Business or Professional Net Income (Less Loss) Amount
+This feature represents the net income from business or professional activities, after accounting for losses.
+It is a measure of the economic success and profitability of businesses within a zip code.
+Higher net incomes suggest prosperous business environments, which can lead to increased vehicle ownership among business owners and employees, reflecting economic prosperity in these areas.
 
-    Individual Retirement Arrangement (IRA) Payments Amount: Contributions to IRAs reflect financial planning and the presence of disposable income.
-Areas with higher IRA payments may have wealthier populations, potentially increasing the likelihood of EV adoption due to greater financial resources.
+9. Real Estate Taxes Amount
+Real estate taxes are directly linked to property values and indirectly to wealth.
+Higher real estate taxes generally indicate more expensive properties, suggesting higher overall wealth in an area.
+Wealthier populations have greater access to credit and financing options, which can increase the likelihood of purchasing vehicles, including newer and more expensive models like EVs.
 
-    Self-Employed (Keogh) Retirement Plans Amount: Similar to IRAs, Keogh plans are retirement savings plans for self-employed individuals.
-Contributions to these plans suggest higher income and financial planning, which can correlate with higher rates of vehicle ownership, including EVs.
-
-    Additional Medicare Tax Amount: This tax applies to high-income earners and indicates a population segment with substantial earnings.
-Higher additional Medicare tax payments correlate with higher income levels, potentially leading to increased EV adoption due to greater financial capacity.
-
-### Implications and Socio-Economic Insights
-
-The analysis of these top features reveals a consistent pattern: higher vehicle counts, particularly for EVs, are associated with higher income levels, wealth indicators (such as mortgage interest and real estate taxes), and financial planning (such as retirement contributions).
-These factors suggest that regions with greater financial capacity are more likely to adopt EVs, reflecting broader trends in consumer behavior and market accessibility.
-
-The emphasis on features related to self-employment and additional taxes for high earners highlights the impact of income diversity on vehicle ownership patterns.
-The presence of deductions and credits for self-employed individuals suggests that entrepreneurial regions may exhibit unique vehicle ownership trends, possibly favoring flexibility and lower-cost vehicles or, conversely, reflecting higher disposable income that supports EV adoption.
+10. Net Premium Tax Credit Amount
+The net premium tax credit amount provides a tax credit under the Affordable Care Act, generally aimed at reducing the cost of health insurance for lower to middle-income families.
+While it may seem counterintuitive, this feature could correlate with regions where financial assistance is necessary, reflecting economic diversity.
+In these areas, the adoption of vehicles, including EVs, may be influenced by financial incentives or subsidies aimed at promoting energy-efficient technologies.
 
@@
 ## Citations
-Jain, A. (n.d.). Electric vehicle population [Data set]. Kaggle. Retrieved June 6, 2024, from https://www.kaggle.com/datasets/jainaru/electric-vehicle-population <br/>
-
-Internal Revenue Service.
-(n.d.).
-SOI tax stats - Individual income tax statistics - ZIP code data (SOI).
-Retrieved June 16, 2024, from https://www.irs.gov/statistics/soi-tax-stats-individual-income-tax-statistics-zip-code-data-soi

### 2024-08-06 products/manuscript/supplement/Supplementary-Material.qmd
add scatterplots and update Manuscript

@@
 ## Log-Transformed Histogram Examples
-![SE Health Insurance Deduction](../../../results/figures/histograms/Self-employed%20health%20insurance%20deduction%20amount_histogram.png)
+![Individual Retirement Payments](../../../results/figures/scatterplots/Individual%20retirement%20arrangement%20payments%20amount_scatterplot.png)
 
@@
 
-![Net Premium Tax Credit Amount](../../../results/figures/histograms/Net%20premium%20tax%20credit%20amount_histogram.png)
+![Home Mortgage Interest Paid](../../../results/figures/histograms/home%20mortgage%20interest%20paid%20amount_histogram.png)
 
 ## Log-Transformed Boxplot Examples
-![SE Health Insurance Deduction](../../../results/figures/boxplots/Self-employed%20health%20insurance%20deduction%20amount_boxplot.png)
+![Individual Retirement Payments](../../../results/figures/boxplots/Individual%20retirement%20arrangement%20payments%20amount_boxplot.png)
 
@@
 
-![Net Premium Tax Credit Amount](../../../results/figures/boxplots/Net%20premium%20tax%20credit%20amount_boxplot.png)
+![Home Mortgage Interest Paid](../../../results/figures/boxplots/home%20mortgage%20interest%20paid%20amount_boxplot.png)
 
@@
 
+## Variable Importance - Decision Tree
+!['Decision Tree Regressor'](../../../results/output/variable_importance/Decision%20Tree%20Regressor.png)
+
+## Variable Importance - Hist Gradient Boosting
+!['Hist Gradient Boosting'](../../../results/output/variable_importance/HistGradientBoostingRegressor%20(Permutation%20Importance).png)
+
+## Variable Importance - Random Forest
+!['Random Forest'](../../../results/output/variable_importance/Random%20Forest%20Regressor.png)
+
 # References

### 2024-08-06 products/manuscript/supplement/Supplementary-Material.qmd
fix graph in supplement material

@@
 ## Log-Transformed Histogram Examples
-![Individual Retirement Payments](../../../results/figures/scatterplots/Individual%20retirement%20arrangement%20payments%20amount_scatterplot.png)
+![Individual Retirement Payments](../../../results/figures/histograms/Individual%20retirement%20arrangement%20payments%20amount_histogram.png)
 

### 2024-08-06 aboutme.qmd
add first site iteration

---
title: "About Me"
---

![The Real Family](IMG_3544.jpg){height=400 fig-alt="The Real Family" fig-align="center"}

#### **Personal Background**
My name is Collin Real. I'm 27 years old, and I grew up in Saint Hedwig, Texas. My hobbies include fitness, technology, and hanging out with my friends and family. I also breed Persian cats.

#### **Education/Career**
I have a Bachelor's and Master's Degree in Accounting, and worked as a Tax Accountant for 3-4 years. As I was finishing my accounting degrees, my friend introduced me to programming, and I instantly knew I was headed for a career change. I researched how Python could be used in accounting/finance, discovering the power of automation, building websites, and data science. To gain experience, I used YouTube, GitHub, and many other websites to learn from projects created by others. After adding this experience to my resume, I joined the Macrospace Analytics team at H-E-B in 2023 as an analyst, performing typical analyst as well as data science work. I hope to transition fully to data science work once I complete the MSDA program.

#### **Storytelling w/ Data**
{{< video https://www.youtube.com/watch?v=8EMW7io4rSI >}}

### 2024-08-06 code/analysis-code/statistical_analysis.qmd
add first site iteration

@@
 features_to_use = data.columns[3:-1]
-print(features_to_use)
 
@@
 
+# Scale data
 scalar = StandardScaler()
 scale = scalar.fit_transform(X_grouped)
-print(scale)
 

### 2024-08-06 code/eda-code/eda-ev-dataset.qmd
add first site iteration

@@
 
-```{r}
-getwd()
-```
+# Introduction
+This document is an exploratory data analysis of electric vehicle data.
+The data was obtained from the IRS and contains information on electric vehicles, their owners, and the tax benefits they receive.
+The data was cleaned and processed in R and saved as a CSV file.
+This document will load the data, explore it, and fit a linear regression model to predict the number of electric vehicles in a given zip code.
 
@@
   filter(Base.MSRP != 0)
-filter_msrp_outlier
 ```
@@
 ```
-
-``` {r}
-# Create a scatterplot - price vs. electric vehicle range correlation
-
-p_scatterplot<-ggplot(filter_msrp_outlier, aes(x = Electric.Range, y = Base.MSRP)) +
-  geom_point() +
-  labs(title = "Scatterplot - Price vs. Range Correlation",
-       x = "Vehicle Range",
-       y = "MSRP")
-plot(p_scatterplot)
-```
-
-```{r}
-#For Electric.Range:
-Q1 <- quantile(ev_data_filtered$Electric.Range , 0.25, na.rm = TRUE) # First Quartile
-Q3 <- quantile(ev_data_filtered$Electric.Range , 0.75, na.rm = TRUE) # Third Quartile
-IQR <- Q3 - Q1 # Compute Interquartile Range
-lower_bound <- Q1 - 1.5 * IQR # Compute Lower Bound based on IQR
-upper_bound <- Q3 + 1.5 * IQR # Compute Upper Bound based on IQR
-upper_bound
-```
-
-```{r}
-# Filter out outliers:
-ev_data_Filtered_2 <- ev_data_filtered[ev_data_filtered$Electric.Range > lower_bound & ev_data_filtered$Electric.Range < upper_bound, ]
-ev_data_Filtered_2
-```

### 2024-08-06 code/eda-code/eda-ev-tax.qmd
add first site iteration

@@
 ---
-title: "EV Exploratory Data Analysis"
+title: "Log-Transform Predictors"
 date: today
@@
 
-# Introduction:
-
-This document is an exploratory data analysis of electric vehicle data.
-The data was obtained from the IRS and contains information on electric vehicles, their owners, and the tax benefits they receive.
-The data was cleaned and processed in R and saved as a CSV file.
-This document will load the data, explore it, and fit a linear regression model to predict the number of electric vehicles in a given zip code.
+# Introduction
+Apply log transformation to our feature
 
@@
 ```
-
+<!-- 
 ```{python}
@@
     plt.scatter(X_grouped[feature], y_grouped, c="lightblue", edgecolors="blue")
-    plt.title(f'{feature} vs Vehicle Count')
+    plt.title(f'{feature} vs. Vehicle Count Relationship')
     plt.xlabel(feature)
@@
 plot_transformed_feature_boxplots(X_grouped, X_transformed)
-```
+``` -->

### 2024-08-06 index.qmd
add first site iteration

--- 2024-08-06 products/manuscript/Manuscript.qmd
@@
 author: Sri Lakshmi Chundru, Collin Real, Joaquin Ramirez, Seth Harris
-format:
-  docx:
-    toc: false
-    number-sections: true
-    highlight-style: github
-bibliography: ../../assets/dataanalysis-references.bib
-csl: ../../assets/apa.csl
+bibliography: assets/dataanalysis-references.bib
+csl: assets/apa.csl
 ---
@@
 #### Scatterplots - Predictor's vs. Vehicle Count Relationship
-![SE Health Insurance Deduction Scatterplot](../../results/figures/scatterplots/Self-employed%20health%20insurance%20deduction%20amount_scatterplot.png)
+![SE Health Insurance Deduction Scatterplot](results/figures/scatterplots/Self-employed%20health%20insurance%20deduction%20amount_scatterplot.png)
 
@@
 #### Log-Transform Histogram
-![SE Health Insurance Deduction](../../results/figures/histograms/Self-employed%20health%20insurance%20deduction%20amount_histogram.png)
+![SE Health Insurance Deduction](results/figures/histograms/Self-employed%20health%20insurance%20deduction%20amount_histogram.png)
 
 #### Log-Transform Boxplot
-![SE Health Insurance Deduction Boxplot](../../results/figures/boxplots/Self-employed%20health%20insurance%20deduction%20amount_boxplot.png)
+![SE Health Insurance Deduction Boxplot](results/figures/boxplots/Self-employed%20health%20insurance%20deduction%20amount_boxplot.png)
 
@@
 <br/>
-![XGBoost](../../results/output/model_results/XGBoost_Regressor.png)
+![XGBoost](results/output/model_results/XGBoost_Regressor.png)
 
@@
 Its R-squared value of 0.857 indicates that it explains 85.7% of the variance, making it the least accurate model in this comparison.
-![Decision Tree](../../results/output/model_results/Decision_Tree_Regressor.png)
+![Decision Tree](results/output/model_results/Decision_Tree_Regressor.png)
 
@@
 While better than the Decision Tree model, it still lagged behind XGBoost and Random Forest in terms of prediction accuracy and variance explanation.
-![HistGradientBoosting](../../results/output/model_results/HistGradientBoosting_Regressor.png)
+![HistGradientBoosting](results/output/model_results/HistGradientBoosting_Regressor.png)
 
@@
 These metrics indicate that the Random Forest model not only provided the most accurate predictions but also explained 88.5% of the variance in vehicle counts, making it the most reliable model in this analysis.
-![Random Forest](../../results/output/model_results/Random_Forest_Regressor.png)
+![Random Forest](results/output/model_results/Random_Forest_Regressor.png)
 
@@
 ## Variable Importance - XGBoost
-!['XGBoost'](../../results/output/variable_importance/XGBoost.png)
+!['XGBoost'](results/output/variable_importance/XGBoost.png)
 

### 2024-08-06 code/analysis-code/statistical_analysis.qmd
add re-render for organization plus minor edits

@@
 title: "Statistical Analysis"
-date: today
 output: html_document
@@
 
-rand_for_feature_importance_list = xgboost_feature_importance['Feature'].head(10).tolist()
-print(rand_for_feature_importance_list)
-
 # Plotting function
 def plot_feature_importance(importances, model_name):
-    plt.figure(figsize=(15, 10))
+    plt.figure(figsize=(10, 5))
     plt.barh(importances['Feature'], importances['Importance'], color='skyblue')
@@
     plt.savefig(f'../../results/output/variable_importance/{model_name}.png')
+    plt.show()
 
@@
 # Display the comparison results
+model_comparison_metrics_df.to_csv('../../results/output/model_comparison.csv', index=False)
+```
+
+### Model Comparison Results
+```{python}
 print(model_comparison_metrics_df)
-model_comparison_metrics_df.to_csv('../../results/output/model_comparison.csv', index=False)
-
 ```

### 2024-08-06 code/eda-code/eda-ev-tax.qmd
add re-render for organization plus minor edits

@@
 title: "Log-Transform Predictors"
-date: today
 output: html_document
@@
 
-# Introduction
-Apply log transformation to our feature
+## Introduction
+Create scatterplots to see the relationship between the independent variable and predictors. Perform a log transformation on features to normalize and compare original to transformed distribution.
 
-# Setup
-
+## Import Libraries & Data, Create Train/Test Datasets
 ```{python}
@@
     X_grouped, y_grouped, test_size=0.2, random_state=42)
+```
 
-merged_data.to_csv('test.csv', index=False)
-```
-<!-- 
+## Scatterplot Relationship - Predictor vs. Independent Variable
 ```{python}
@@
     plt.ylabel('Vehicle Count')
-    # plt.grid(True)
     plt.savefig(f'../../results/figures/scatterplots/{feature}_scatterplot.png')
-
+    plt.show()
 ```
 
+## Histogram Distributions
 ```{python}
@@
         ax1.title.set_text(f'Original vs. Log-Transformed: {feature}')
-        # ax2.title.set_text(f'Log-Transformed Histogram of {feature}')
         plt.xlabel(feature)
@@
         plt.grid(True)
+        plt.tight_layout()
         plt.savefig(f'../../results/figures/histograms/{feature}_histogram.png')
+        plt.show()
 
@@
 
+## Boxplot Distributions
 ```{python}
@@
         ax1.title.set_text(f'Original vs. Log-Transformed Boxplot of {feature}')
-        # ax2.title.set_text(f'Log-Transformed Boxplot of {feature}')
         plt.xlabel(feature)
@@
         plt.savefig(f'../../results/figures/boxplots/{feature}_boxplot.png')
+        plt.show()
 
@@
 plot_transformed_feature_boxplots(X_grouped, X_transformed)
-``` -->
+```

### 2024-08-06 index.qmd
add re-render for organization plus minor edits

@@
 ---
-title: "Using IRS Data to Predict Electric Vehicle Count"
+title: "EV Predictive Modeling"
+subtitle: "IRS Tax Data - Zip Code"
 author: Sri Lakshmi Chundru, Collin Real, Joaquin Ramirez, Seth Harris
@@
 
-1. Self-Employed Health Insurance Deduction Amount
+1. Self-Employed Health Insurance Deduction Amount:
 This feature represents the amount self-employed individuals can deduct for health insurance premiums.
@@
 
-2. Individual Retirement Arrangement Payments Amount
-Contributions to Individual Retirement Arrangements (IRAs) reflect financial planning and savings behavior.
+2.
+Individual Retirement Arrangement Payments Amount Contributions to Individual Retirement Arrangements (IRAs) reflect financial planning and savings behavior.
 Higher contributions suggest higher disposable income and financial stability, factors that are positively associated with the ability to purchase vehicles, including more expensive options like EVs.
@@
 
-3. Home Mortgage from Personal Seller Amount
+3. Home Mortgage from Personal Seller Amount:
 This feature involves transactions where the home mortgage is obtained directly from a personal seller rather than a traditional financial institution.
@@
 
-4. Salaries and Wages Amount
+4. Salaries and Wages Amount:
 A direct indicator of income, this feature represents the total salaries and wages reported within a zip code.
@@
 
-5. Home Mortgage Interest Paid Amount
+5. Home Mortgage Interest Paid Amount:
 The amount of interest paid on home mortgages is indicative of both property ownership and the financial burden of maintaining such properties.
@@
 
-6. Self-Employment Tax Amount
+6. Self-Employment Tax Amount:
 This feature captures the amount of self-employment tax paid, which directly correlates with self-employment income.
@@
 
-7. Self-Employed (Keogh) Retirement Plans Amount
-Contributions to Keogh plans reflect savings set aside by self-employed individuals for retirement.
+7.
+Self-Employed (Keogh) Retirement Plans Amount Contributions to Keogh plans reflect savings set aside by self-employed individuals for retirement.
 Like IRA contributions, this feature suggests a focus on long-term financial security, indicative of higher income levels and financial planning.
@@
 
-8. Business or Professional Net Income (Less Loss) Amount
+8. Business or Professional Net Income (Less Loss) Amount:
 This feature represents the net income from business or professional activities, after accounting for losses.
@@
 
-9. Real Estate Taxes Amount
+9. Real Estate Taxes Amount:
 Real estate taxes are directly linked to property values and indirectly to wealth.
@@
 
-10. Net Premium Tax Credit Amount
+10. Net Premium Tax Credit Amount:
 The net premium tax credit amount provides a tax credit under the Affordable Care Act, generally aimed at reducing the cost of health insurance for lower to middle-income families.
